---
title: "1. Gapfilling Council from Kyle 7.16.2024" #Gapfill the met data for Council 
#Here we combined our site data with the ERA5 data to check agreement, used the linear regression to adjust the ERA5 model to make it match even better, and now we are merging that back in with our dataset to use for gapfilling 

#11.25.2024: adding in u and v components for wind direction 
output: html_document
date: "2024-09-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = 'C:/Users/karndt.WHRC/Desktop/sites/YKD/Ameriflux YK/unburned/') #use this to set your directory or use the code in the next chunk to set working directory 
```

#Libraries & Data 
```{r}
rm(list = ls())

library(data.table)
library(ggplot2)
library(zoo)
library(cowplot)
library(dplyr)
Sys.setenv(TZ = "UTC")

#set working directory
setwd("C:/Users/kkent/Documents/Council Data/Council BASE gapfilling")
 
#double check working directory 
getwd()

#Load your "cleaned" met dataframe
df = fread(input = "C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/councilbase_met_flux_clean_2017_2023.csv",na.strings = c('-9999'))#redoing with re-cleaned data from Dani & with wind direction 

#making all columns numeric, they read in as.character, need to change to as.numeric to have functions work 
df <- df %>%
  mutate(across(where(is.character), as.numeric))

```

#Load Data and adjust timestamp to set up POSIXct
```{r}
#Breaks out the timing by yr, month, day, hour, min, sec, and sets 0 values to -9999

#formatting the timestamp 
year = substr(x = df$TIMESTAMP_END,start = 1,stop = 4)
month = substr(x = df$TIMESTAMP_END,start = 5,stop = 6)
day = substr(x = df$TIMESTAMP_END,start = 7,stop = 8)
hour = substr(x = df$TIMESTAMP_END,start = 9,stop = 10)
min = substr(x = df$TIMESTAMP_END,start = 11,stop = 12)

date = paste(year,month,day,sep = '-')
time = paste(hour,min,sep = ':')
ts = paste(date,time,sep = ' ')

df$ts = as.POSIXct(ts)

#check timestamp - use a random range of dates to make sure the timestamp is correctly mapped. Here we map shortwave_incoming by day so you should see the diurnal pattern across days 
ggplot(data = df)+
  geom_point(aes(ts,SW_IN.c))+
  scale_x_datetime(limits = as.POSIXct(c("2019-06-06","2019-06-10")))

#remove duplicated time stamp rows
df = df[!duplicated(df$ts),]
```
gap fill short gaps in the met data first
```{r}

#10.28.2024 - KK - added in ".c" behind re-cleaned met data from Dani

gap = 3 #represents an hour and a half (3 30min sections), helps fill little data-logger gaps/issues

#energy fluxes ###########################
df$G_1_1_1.c = na.spline(df$G_1_1_1.c,maxgap = gap)

#meteorology and weather #################################
df$TA.c  = na.spline(df$TA.c,maxgap = gap)
df$RH.c  = na.spline(df$RH.c,maxgap = gap)

#Radiation #####################################
df$LW_IN.c  = na.spline(df$LW_IN.c,maxgap = gap)
df$LW_OUT.c = na.spline(df$LW_OUT.c,maxgap = gap)
df$SW_IN.c  = na.spline(df$SW_IN.c,maxgap = gap)
df$SW_OUT.c = na.spline(df$SW_OUT.c,maxgap = gap)

df$NETRAD.c = na.spline(df$NETRAD.c,maxgap = gap)

df$PPFD_IN.c = na.spline(df$PPFD_IN.c,maxgap = gap)
df$PPFD_OUT.c = na.spline(df$PPFD_OUT.c,maxgap = gap)

#Soil factors ###################################
#all swc in %
df$SWC_1_1_1.c = na.spline(df$SWC_1_1_1.c,maxgap = gap)
df$SWC_2_1_1.c = na.spline(df$SWC_2_1_1.c,maxgap = gap)
df$SWC_3_1_1.c = na.spline(df$SWC_3_1_1.c,maxgap = gap)

#all soil temps deg C
df$TS_1_1_1.c = na.spline(df$TS_1_1_1.c,maxgap = gap)
df$TS_2_1_1.c = na.spline(df$TS_2_1_1.c,maxgap = gap)
df$TS_3_1_1.c = na.spline(df$TS_3_1_1.c,maxgap = gap)
df$TS_4_1_1.c = na.spline(df$TS_4_1_1.c,maxgap = gap)

#can't gapfill WD as it makes the degrees -50 - 550, and BASE has no u and v components 

```


#Add cardinal directions for wind direction (WD) in BASE df
```{r}
#make sure windDir is numeric 

#(0=north,90=east,180=south,270=west) that the wind is coming from
# northerly wind is 0°, an easterly wind is 90°, a southerly wind is 180°, and a westerly wind is 270°

# Function to assign cardinal directions
df <- df %>%
  mutate(
    Cardinal_Direction = case_when(
      WD >= 337.5 | WD < 22.5  ~ "N",
      WD >= 22.5 & WD < 67.5  ~ "NE",
      WD >= 67.5 & WD < 112.5 ~ "E",
      WD >= 112.5 & WD < 157.5 ~ "SE",
      WD >= 157.5 & WD < 202.5 ~ "S",
      WD >= 202.5 & WD < 247.5 ~ "SW",
      WD >= 247.5 & WD < 292.5 ~ "W",
      WD >= 292.5 & WD < 337.5 ~ "NW",
      TRUE ~ NA_character_ # Handle any missing or invalid values -- not needed in this case 
    )
  )

#checking to see if NA  match up correctly  
sum(is.na(df$WD)) #66561
sum(is.na(df$Cardinal_Direction)) #66561

```


#Save new file of the cleaned met data with minor gapfilling 
```{r}
write.csv(x = df,file = "C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_2016_2023.csv",row.names = F)
```

#Load in and Add ERA5 data
```{r}
#load the ERA5 df from step 0 --> edited to include u and v wind components 

era = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_era5.csv')

df$date = df$ts

summary(era$date) #checking out the timestamps 

#merge cleaned council met df and ERA5 df together
all = merge(era,df,by = 'date',all = T)
all = all[!duplicated(all$date),]
```
#Here we use linear models to make adjustments to the ERA5 data to make it better fit our actual site data and improve later gapfilling


#checking data, and running linear regressions to be able to adjust the ERA5 data to better represent what data from the site is showing*
#the linear reg applies an adjustment to the ERA5 data 
#green data is overwriting the ERA5 pink data -- and we use that linear-reg-adjusted data to help fill in the site data as it's a better representation*

#wind direction in degrees (0-360)
```{r}
#lowercase variables come from ERA; upper-case from our cleaned council met dataset
#KK 10.28.2024: added wind direction 
#from Kyle - these might not agree well as they could come from different height profiles...so while they don't seem to align well at all here, let's not worry about this for now. 

wd.mod = lm(formula = all$WD ~ all$wd)
summary(wd.mod)

all$wd.eramod = all$wd*wd.mod$coefficients[2] + wd.mod$coefficients[1]   

ggplot(data = all)+
  geom_point(aes(date,wd,col='ERA5'))+
  geom_point(aes(date,WD,col='Site'))

ggplot(data = all,aes(wd,WD))+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)
```

#RH - relative humidity 
```{r}
#lowercase variables come from ERA; upper-case from our cleaned council met dataset
#KK 10.28.2024: added the ".c" to use re-cleaned data from Dani 

rh.mod = lm(formula = all$RH.c ~ all$rh)
summary(rh.mod) #R2 = 0.23, slope = 0.76

all$rh.eramod = all$rh*rh.mod$coefficients[2] + rh.mod$coefficients[1]   

ggplot(data = all)+
  geom_point(aes(date,rh,col='ERA5'))+
  geom_point(aes(date,RH.c,col='Site'))

ggplot(data = all,aes(rh,RH.c))+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)

#make cleaner fig to show ERA5 vs site measurements for paper 

RH_plot <- ggplot(data = all, aes(x = rh, y = RH.c)) +
  theme_bw() +
  geom_point() +
  # Linear fit (blue solid line)
  geom_smooth(method = "lm",
              aes(color = "Linear fit", linetype = "Linear fit"),
              se = FALSE) +
  # 1:1 line (red dashed line)
  geom_abline(aes(slope = 1, intercept = 0,
                  color = "1:1 line", linetype = "1:1 line")) +
  # Axis labels
  labs(x = "ERA5 Relative Humidity (%)",
       y = "Site Relative Humidity (%)",
       color = "", linetype = "") +
  # Manually set colors and linetypes
  scale_color_manual(values = c("Linear fit" = "blue", "1:1 line" = "red")) +
  scale_linetype_manual(values = c("Linear fit" = "solid", "1:1 line" = "dashed")) +
  theme(legend.position = "bottom") +
  # Annotations
  annotate("text",
           x = 30, y = 140,
           hjust = -0.1, vjust = 1.3,
           label = "R² = 0.23", size = 5) +
  annotate("text",
           x = 30, y = 125,
           hjust = -0.1, vjust = 1.3,
           label = "Slope = 0.76", size = 5) +
    #Title for the plot 
   labs(title = "Relative Humidity") 

print(RH_plot)


#Modified ERA5 vs site agreement 
rh_modified = lm(formula = all$RH.c ~ all$rh.eramod)
summary(rh_modified) #R2 = 0.23, slope = 1

RH_plot2 <- ggplot(data = all, aes(x = rh.eramod, y = RH.c)) +
  theme_bw() +
  geom_point() +
  # Linear fit (blue solid line)
  geom_smooth(method = "lm",
              aes(color = "Linear fit", linetype = "Linear fit"),
              se = FALSE) +
  # 1:1 line (red dashed line)
  geom_abline(aes(slope = 1, intercept = 0,
                  color = "1:1 line", linetype = "1:1 line")) +
  # Axis labels
  labs(x = "Modified ERA5 Relative Humidity (%)",
       y = "Site Relative Humidity (%)",
       color = "", linetype = "") +
  # Manually set colors and linetypes
  scale_color_manual(values = c("Linear fit" = "blue", "1:1 line" = "red")) +
  scale_linetype_manual(values = c("Linear fit" = "solid", "1:1 line" = "dashed")) +
  theme(legend.position = "bottom") +
  # Annotations
  annotate("text",
           x = 20, y = 140,
           hjust = -0.1, vjust = 1.3,
           label = "R² = 0.23", size = 5) +
  annotate("text",
           x = 20, y = 125,
           hjust = -0.1, vjust = 1.3,
           label = "Slope = 1", size = 5) +
    #Title for the plot 
   labs(title = "Modified ERA5 Relative Humidity") 

print(RH_plot2)

```
#AirT
```{r}
ggplot(data = all)+
  geom_point(aes(date,airt,col='ERA5'))+
  geom_point(aes(date,all$TA.c,col='Site'))

airt.mod = lm(formula = all$TA.c ~ all$airt)
summary(airt.mod) #R2 = 0.96; slope = 0.94

all$airt.eramod = all$airt*airt.mod$coefficients[2] + airt.mod$coefficients[1]     

ggplot(data = all,aes(airt,TA.c))+theme_bw()+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)


#make cleaner fig to show ERA5 and site air temp agreement for paper

airtemp_plot <- ggplot(data = all, aes(x = airt, y = TA.c)) +
  theme_bw() +
  geom_point() +
  # Linear fit (blue solid line)
  geom_smooth(method = "lm",
              aes(color = "Linear fit", linetype = "Linear fit"),
              se = FALSE) +
  # 1:1 line (red dashed line)
  geom_abline(aes(slope = 1, intercept = 0,
                  color = "1:1 line", linetype = "1:1 line")) +
  # Axis labels
  labs(x = "ERA5 Air Temperature (°C)",
       y = "Site Air Temperature (°C)",
       color = "", linetype = "") +
  # Manually set colors and linetypes
  scale_color_manual(values = c("Linear fit" = "blue", "1:1 line" = "red")) +
  scale_linetype_manual(values = c("Linear fit" = "solid", "1:1 line" = "dashed")) +
  theme(legend.position = "bottom") +
  # Annotations
  annotate("text",
           x = -41, y = 30,
           hjust = -0.1, vjust = 1.3,
           label = "R² = 0.96", size = 5) +
  annotate("text",
           x = -41, y = 25,
           hjust = -0.1, vjust = 1.3,
           label = "Slope = 0.94", size = 5) +
    #Title for the plot 
   labs(title = "Air Temperature") 

print(airtemp_plot)


#Modified ERA5 vs site agreement 
airt_modified = lm(formula = all$TA.c ~ all$airt.eramod)
summary(airt_modified) #R2 = 0.96; slope = 1

airtemp_plot2 <- ggplot(data = all, aes(x = airt.eramod, y = TA.c)) +
  theme_bw() +
  geom_point() +
  # Linear fit (blue solid line)
  geom_smooth(method = "lm",
              aes(color = "Linear fit", linetype = "Linear fit"),
              se = FALSE) +
  # 1:1 line (red dashed line)
  geom_abline(aes(slope = 1, intercept = 0,
                  color = "1:1 line", linetype = "1:1 line")) +
  # Axis labels
  labs(x = "Modified ERA5 Air Temperature (°C)",
       y = "Site Air Temperature (°C)",
       color = "", linetype = "") +
  # Manually set colors and linetypes
  scale_color_manual(values = c("Linear fit" = "blue", "1:1 line" = "red")) +
  scale_linetype_manual(values = c("Linear fit" = "solid", "1:1 line" = "dashed")) +
  theme(legend.position = "bottom") +
  # Annotations
  annotate("text",
           x = -41, y = 30,
           hjust = -0.1, vjust = 1.3,
           label = "R² = 0.96", size = 5) +
  annotate("text",
           x = -41, y = 25,
           hjust = -0.1, vjust = 1.3,
           label = "Slope = 0.1", size = 5) +
    #Title for the plot 
   labs(title = "Modified ERA5 Air Temperature") 

print(airtemp_plot2)



```


#SoilT
```{r}
ggplot(data = all)+
  geom_point(aes(date,st1,col='ERA5.1'))+
  geom_point(aes(date,st2,col='ERA5.2'))+
  geom_point(aes(date,all$TS_1_1_1.c,col='Site1'))+
  geom_point(aes(date,all$TS_2_1_1.c,col='Site2'))+
  geom_point(aes(date,all$TS_3_1_1.c,col='Site3'))+
   geom_point(aes(date,all$TS_4_1_1.c,col='Site4'))
   
#changed from TS2 to TS3 (tussock) & re-ran code on 6/16/2025
soilt.mod = lm(formula = all$TS_3_1_1.c ~ all$st2)
summary(soilt.mod) #R2 = 0.87, slope = 0.797 = 0.80

all$tsoil.eramod = all$st2*soilt.mod$coefficients[2] + soilt.mod$coefficients[1]           

ggplot(data = all)+
  geom_point(aes(date,st2,col='ERA5'))+
  geom_point(aes(date,TS_3_1_1.c,col='Site'))+
  geom_point(aes(date,tsoil.eramod,col='Lin. Reg.'))

#make cleaner fig to show ERA5 and site soil temp agreement for paper

soiltemp_plot <- ggplot(data = all, aes(x = st2, y = TS_3_1_1.c)) +
  theme_bw() +
  geom_point() +
  # Linear fit (blue solid line)
  geom_smooth(method = "lm",
              aes(color = "Linear fit", linetype = "Linear fit"),
              se = FALSE) +
  # 1:1 line (red dashed line)
  geom_abline(aes(slope = 1, intercept = 0,
                  color = "1:1 line", linetype = "1:1 line")) +
  # Axis labels
  labs(x = "ERA5 Soil Temperature (°C)",
       y = "Site Soil Temperature at 15cm (°C)",
       color = "", linetype = "") +
  # Manually set colors and linetypes
  scale_color_manual(values = c("Linear fit" = "blue", "1:1 line" = "red")) +
  scale_linetype_manual(values = c("Linear fit" = "solid", "1:1 line" = "dashed")) +
  theme(legend.position = "bottom") +
  # Annotations
  annotate("text",
           x = -10, y = 30,
           hjust = -0.1, vjust = 1.3,
           label = "R² = 0.87", size = 5) +
  annotate("text",
           x = -10, y = 25,
           hjust = -0.1, vjust = 1.3,
           label = "Slope = 0.80", size = 5) +
    #Title for the plot 
   labs(title = "Soil Temperature") 

print(soiltemp_plot)


#modified ERA5
soilt_modified = lm(formula = all$TS_3_1_1.c ~ all$tsoil.eramod)
summary(soilt_modified) #R2 = 0.87, slope = 1


# # Fill gaps using the linear model
# df$TS_3_1_1_gapfilled[gaps] <- predict(gapfill_model, 
#                                         newdata = df[gaps, ])


```

#Rg (SW-IN)  - global radiation 
```{r}
# "rad" is from ERA5; SW-IN is from our re-cleaned data, so changed to SW-IN.c to use the re-cleaned data 

ggplot(data = all)+
  geom_point(aes(date,rad,col='ERA5'))+
  geom_point(aes(date,all$SW_IN.c,col='Site'))

ggplot(data = all)+
  geom_point(aes(date,rad,col='ERA5'))+
  geom_point(aes(date,SW_IN.c,col='Site'))+
  scale_x_datetime(limits = as.POSIXct(c('2019-06-20','2019-07-1')))

rad.mod = lm(formula = all$SW_IN.c ~ all$rad)
summary(rad.mod) #R2 = 0.8459; slope = 0.98

all$rad.eramod = all$rad*rad.mod$coefficients[2] + rad.mod$coefficients[1]   

ggplot(data = all,aes(rad,SW_IN.c))+theme_bw()+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)


#make cleaner fig to show ERA5 vs site agreement for paper 

rad_plot <- ggplot(data = all, aes(x = rad, y = SW_IN.c)) +
  theme_bw() +
  geom_point() +
  # Linear fit (blue solid line)
  geom_smooth(method = "lm",
              aes(color = "Linear fit", linetype = "Linear fit"),
              se = FALSE) +
  # 1:1 line (red dashed line)
  geom_abline(aes(slope = 1, intercept = 0,
                  color = "1:1 line", linetype = "1:1 line")) +
  # Axis labels
  labs(x = "ERA5 Radiation (W/m²)",
       y = "Site SW_IN (W/m²)",
       color = "", linetype = "") +
  # Manually set colors and linetypes
  scale_color_manual(values = c("Linear fit" = "blue", "1:1 line" = "red")) +
  scale_linetype_manual(values = c("Linear fit" = "solid", "1:1 line" = "dashed")) +
  theme(legend.position = "bottom") +
  # Annotations
  annotate("text",
           x = 0, y = 990,
           hjust = -0.1, vjust = 1.3,
           label = "R² = 0.84", size = 5) +
  annotate("text",
           x = 0, y = 930,
           hjust = -0.1, vjust = 1.3,
           label = "Slope = 0.98", size = 5) +
    #Title for the plot 
   labs(title = "Radiation") 

print(rad_plot)


#modified ERA5
rad_modified = lm(formula = all$SW_IN.c ~ all$rad.eramod)
summary(rad_modified) #R2 = 0.84, slope = 1

```


#ws - wind speed 
```{r}
#WS not re-cleaned, so keeping the original WS variable  
ggplot(data = all)+
  geom_point(aes(date,ws,col='ERA5'))+
  geom_point(aes(date,all$WS,col='Site'))

ws.mod = lm(formula = all$WS ~ all$ws)
summary(ws.mod) #R2 = 0.45; slope = 0.898

all$ws.eramod = all$ws*ws.mod$coefficients[2] + ws.mod$coefficients[1]    

ggplot(data = all,aes(ws,WS))+theme_bw()+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)

#make cleaner fig to show ERA5 vs site agreement for paper 

ws_plot <- ggplot(data = all, aes(x = ws, y = WS)) +
  theme_bw() +
  geom_point() +
  # Linear fit (blue solid line)
  geom_smooth(method = "lm",
              aes(color = "Linear fit", linetype = "Linear fit"),
              se = FALSE) +
  # 1:1 line (red dashed line)
  geom_abline(aes(slope = 1, intercept = 0,
                  color = "1:1 line", linetype = "1:1 line")) +
  # Axis labels
  labs(x = "ERA5 WS (m/s)",
       y = "Site WS (m/s)",
       color = "", linetype = "") +
  # Manually set colors and linetypes
  scale_color_manual(values = c("Linear fit" = "blue", "1:1 line" = "red")) +
  scale_linetype_manual(values = c("Linear fit" = "solid", "1:1 line" = "dashed")) +
  theme(legend.position = "bottom") +
  # Annotations
  annotate("text",
           x = 0, y = 20,
           hjust = -0.1, vjust = 1.3,
           label = "R² = 0.45", size = 5) +
  annotate("text",
           x = 0, y = 18,
           hjust = -0.1, vjust = 1.3,
           label = "Slope = 0.90", size = 5) +
    #Title for the plot 
   labs(title = "Wind Speed") 

print(ws_plot)


#modified ERA5
ws_modified = lm(formula = all$WS ~ all$ws.eramod)
summary(ws_modified) #R2 = 0.45, slope = 1

```


#Le - latent heat flux 
```{r}
#changed to LE.c to use re-cleaned data 
ggplot(data = all)+
  geom_point(aes(date,le,col='ERA5'))+
  geom_point(aes(date,LE.c,col='Site'))

le.mod = lm(formula = all$LE.c ~ all$le)
summary(le.mod) #R2 = 0.55, slope = 0.55

all$le.eramod = all$le*le.mod$coefficients[2] + le.mod$coefficients[1]    

ggplot(data = all,aes(le,LE.c))+theme_bw()+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)


#make cleaner fig to show ERA5 vs site agreement for paper 

LE_plot <- ggplot(data = all, aes(x = le, y = LE.c)) +
  theme_bw() +
  geom_point() +
  # Linear fit (blue solid line)
  geom_smooth(method = "lm",
              aes(color = "Linear fit", linetype = "Linear fit"),
              se = FALSE) +
  # 1:1 line (red dashed line)
  geom_abline(aes(slope = 1, intercept = 0,
                  color = "1:1 line", linetype = "1:1 line")) +
  # Axis labels
  labs(x = "ERA5 LE (W/m²)",
       y = "Site LE (W/m²)",
       color = "", linetype = "") +
  # Manually set colors and linetypes
  scale_color_manual(values = c("Linear fit" = "blue", "1:1 line" = "red")) +
  scale_linetype_manual(values = c("Linear fit" = "solid", "1:1 line" = "dashed")) +
  theme(legend.position = "bottom") +
  # Annotations
  annotate("text",
           x = -50, y = 450,
           hjust = -0.1, vjust = 1.3,
           label = "R² = 0.55", size = 5) +
  annotate("text",
           x = -50, y = 400,
           hjust = -0.1, vjust = 1.3,
           label = "Slope = 0.55", size = 5) +
    #Title for the plot 
   labs(title = "Latent Heat Turbulent Flux") 

print(LE_plot)


#modified ERA5
LE_modified = lm(formula = all$LE.c ~ all$le.eramod)
summary(LE_modified) #R2 = 0.55, slope = 1


```
#H - sensible heat flux 
```{r}
#changed to H.c to use re-cleaned data 
ggplot(data = all)+
  geom_point(aes(date,h,col='ERA5'))+
  geom_point(aes(date,H.c,col='Site'))

h.mod = lm(formula = all$H.c ~ all$h)
summary(h.mod)#R2 = 0.6552, slope = 0.84

all$h.eramod = all$h*h.mod$coefficients[2] + h.mod$coefficients[1]    

ggplot(data = all,aes(h,H.c))+theme_bw()+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)


#make cleaner fig to show ERA5 vs site agreement for paper 

H_plot <- ggplot(data = all, aes(x = h, y = H.c)) +
  theme_bw() +
  geom_point() +
  # Linear fit (blue solid line)
  geom_smooth(method = "lm",
              aes(color = "Linear fit", linetype = "Linear fit"),
              se = FALSE) +
  # 1:1 line (red dashed line)
  geom_abline(aes(slope = 1, intercept = 0,
                  color = "1:1 line", linetype = "1:1 line")) +
  # Axis labels
  labs(x = "ERA5 H (W/m²)",
       y = "Site H (W/m²)",
       color = "", linetype = "") +
  # Manually set colors and linetypes
  scale_color_manual(values = c("Linear fit" = "blue", "1:1 line" = "red")) +
  scale_linetype_manual(values = c("Linear fit" = "solid", "1:1 line" = "dashed")) +
  theme(legend.position = "bottom") +
  # Annotations
  annotate("text",
           x = -150, y = 400,
           hjust = -0.1, vjust = 1.3,
           label = "R² = 0.65", size = 5) +
  annotate("text",
           x = -150, y = 350,
           hjust = -0.1, vjust = 1.3,
           label = "Slope = 0.84", size = 5) +
    #Title for the plot 
   labs(title = "Sensible Heat Turbulent Flux") 

print(H_plot)


#modified ERA5
H_modified = lm(formula = all$H.c ~ all$h.eramod)
summary(H_modified) #R2 = 0.65, slope = 1

```
#arrange ERA5 vs site plots into single figure to show comparisons for paper
```{r}
# install.packages("patchwork")  # if needed
library(patchwork)
library(ggplot2)

# Arrange (3x3-ish) and collect a single legend at the bottom
combined <-
  (RH_plot | airtemp_plot | soiltemp_plot) /
  (rad_plot | ws_plot | LE_plot) /
  (H_plot   | plot_spacer() | plot_spacer()) +
  plot_layout(guides = "collect") &                 # share one legend
  theme(legend.position = "bottom")                 # put the legend below

# Add a single title
final_plot <- combined + plot_annotation(
  title = "ERA5 vs Site Comparisons",
  theme = theme(plot.title = element_text(hjust = 0.5, face = "bold"))
)

#make legend bigger 
final_plot <- final_plot &  # Increase legend text size, key size, and spacing
  theme(
    #legend.title = element_text(size = 14),    # Increase legend title size
    legend.text = element_text(size = 14, face = "bold"),     # Increase legend text size
    legend.key.size = unit(2, "cm")) 

final_plot

# Then save it
 ggsave("C:/Users/kkent/Documents/Github Flux Network/Council_Flux_Analysis_Paper/Council Figures/combined_ERA5vsSite_met_variables.png", final_plot, width = 20, height = 20, dpi = 600)

```


```{r}
#remove duplicated timestamp rows
df = df[!duplicated(df$ts),]
```


```{r}
#Added ".c" to cleaned data 

#calculate VPD (vapor pressure deficit) from air t and RH in ERA5 df
svp = 610.7*10^((7.5*all$airt)/(237.3+all$airt))
all$vpd.era = ((100 - all$rh)/100)*svp  

#calculate VPD from air t and RH from cleaned council met df (with ".c" to use re-cleaned data)
svp = 610.7*10^((7.5*all$TA.c)/(237.3+all$TA.c))
all$vpd.site = ((100 - all$RH.c)/100)*svp 


ta = ggplot(data = all,aes(airt,TA.c))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 Air T. ("*degree*"C)"))+
  scale_y_continuous(expression("Site Air T. ("*degree*"C)"))

rh = ggplot(data = all,aes(rh,RH.c))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 RH (%)"))+
  scale_y_continuous(expression("Site RH (%)"))

vpd = ggplot(data = all,aes(vpd.era,vpd.site))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 VPD ("*Wm^-2*")"))+
  scale_y_continuous(expression("Site VPD ("*Wm^-2*")"))

#changed from TS2 to TS3 6/16/25 & re-ran code 
st = ggplot(data = all,aes(st2,TS_3_1_1.c))+theme_bw()+
    geom_point(alpha = 0.25)+
    geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 Soil T. ("*degree*"C)"))+
  scale_y_continuous(expression("Site Soil T. ("*degree*"C)"))

sw = ggplot(data = all,aes(rad,SW_IN.c))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 SW ("*Wm^-2*")"))+
  scale_y_continuous(expression("Site SW ("*Wm^-2*")"))

ws = ggplot(data = all,aes(ws,WS))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 WS ("*ms^-1*")"))+
  scale_y_continuous(expression("Site WS ("*ms^-1*")"))

#added wind direction
wd = ggplot(data = all,aes(wd,WD))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 WD ("*ms^-1*")"))+
  scale_y_continuous(expression("Site WD ("*ms^-1*")"))

le = ggplot(data = all,aes(le,LE.c))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 LE ("*Wm^-2*")"))+
  scale_y_continuous(expression("Site LE ("*Wm^-2*")"))

h = ggplot(data = all,aes(h,H.c))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 H ("*Wm^-2*")"))+
  scale_y_continuous(expression("Site H ("*Wm^-2*")"))

#changed from SWC1 to SWC3 6/16/2025
swc = ggplot(data = all,aes(vwc1*100,SWC_3_1_1.c))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 VWC (%)"))+
  scale_y_continuous(expression("Site VWC (%)"))

#Plot agreement between ERA5 and site data 
#png(filename = 'C:/Users/karndt.WHRC/Desktop/sites/YKD/plots/unburnedera5.png',width = 8,height = 8,units = 'in',res = 1500)
plot_grid(ta,rh,vpd,st,sw,ws,wd,le,h,swc)
#dev.off()

#All look decent except for SWC and WD which clearly are having a rough time

```
#Select variables of interest from ERA5 / Reduce down to Council size; add adjusted ERA5 model to final data set
```{r}
#select the data / variables you want from the ERA5 instead of saving all
eras = all[,c('date','rh.eramod','airt.eramod','ws.eramod','wd', 'cardinal_direction','u', 'v','rad.eramod','tsoil.eramod','le.eramod','h.eramod')] #adding the u, v, wd (not wd.eramod since it wasn't aligning well at all), and cardinal_direction 

df = df[!duplicated(df$date),] #delete duplicated dates
df = merge(df,eras,by = 'date',all.x = T) #merge all of those ERA5 variables into our Council dataset 

write.csv(x = df,file = 'C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_2016_2023_era.csv',row.names = F,quote = F)
```




