---
title: "1. Gapfilling Council from Kyle 7.16.2024" #Gapfill the met data for Council 
#Here we combined our site data with the ERA5 data to check agreement, used the linear regression to adjust the ERA5 model to make it match even better, and now we are merging that back in with our dataset to use for gapfilling 

#11.25.2024: adding in u and v components for wind direction 
output: html_document
date: "2024-09-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = 'C:/Users/karndt.WHRC/Desktop/sites/YKD/Ameriflux YK/unburned/') #use this to set your directory or use the code in the next chunk to set working directory 
```

#Libraries & Data 
```{r}
rm(list = ls())

library(data.table)
library(ggplot2)
library(zoo)
library(cowplot)
library(dplyr)
Sys.setenv(TZ = "UTC")

#set working directory
setwd("C:/Users/kkent/Documents/Council Data/Council BASE gapfilling")
 
#double check working directory 
getwd()

#Load your "cleaned" met dataframe
df = fread(input = "C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/councilbase_met_flux_clean_2017_2023.csv",na.strings = c('-9999'))#redoing with re-cleaned data from Dani & with wind direction 

#making all columns numeric, they read in as.character, need to change to as.numeric to have functions work 
df <- df %>%
  mutate(across(where(is.character), as.numeric))

```

#Load Data and adjust timestamp to set up POSIXct
```{r}
#Breaks out the timing by yr, month, day, hour, min, sec, and sets 0 values to -9999

#formatting the timestamp 
year = substr(x = df$TIMESTAMP_END,start = 1,stop = 4)
month = substr(x = df$TIMESTAMP_END,start = 5,stop = 6)
day = substr(x = df$TIMESTAMP_END,start = 7,stop = 8)
hour = substr(x = df$TIMESTAMP_END,start = 9,stop = 10)
min = substr(x = df$TIMESTAMP_END,start = 11,stop = 12)

date = paste(year,month,day,sep = '-')
time = paste(hour,min,sep = ':')
ts = paste(date,time,sep = ' ')

df$ts = as.POSIXct(ts)

#check timestamp - use a random range of dates to make sure the timestamp is correctly mapped. Here we map shortwave_incoming by day so you should see the diurnal pattern across days 
ggplot(data = df)+
  geom_point(aes(ts,SW_IN.c))+
  scale_x_datetime(limits = as.POSIXct(c("2019-06-06","2019-06-10")))

#remove duplicated time stamp rows
df = df[!duplicated(df$ts),]
```
gap fill short gaps in the met data first
```{r}

#10.28.2024 - KK - added in ".c" behind re-cleaned met data from Dani

gap = 3 #represents an hour and a half (3 30min sections), helps fill little data-logger gaps/issues

#energy fluxes ###########################
df$G_1_1_1.c = na.spline(df$G_1_1_1.c,maxgap = gap)

#meteorology and weather #################################
df$TA.c  = na.spline(df$TA.c,maxgap = gap)
df$RH.c  = na.spline(df$RH.c,maxgap = gap)

#Radiation #####################################
df$LW_IN.c  = na.spline(df$LW_IN.c,maxgap = gap)
df$LW_OUT.c = na.spline(df$LW_OUT.c,maxgap = gap)
df$SW_IN.c  = na.spline(df$SW_IN.c,maxgap = gap)
df$SW_OUT.c = na.spline(df$SW_OUT.c,maxgap = gap)

df$NETRAD.c = na.spline(df$NETRAD.c,maxgap = gap)

df$PPFD_IN.c = na.spline(df$PPFD_IN.c,maxgap = gap)
df$PPFD_OUT.c = na.spline(df$PPFD_OUT.c,maxgap = gap)

#Soil factors ###################################
#all swc in %
df$SWC_1_1_1.c = na.spline(df$SWC_1_1_1.c,maxgap = gap)
df$SWC_2_1_1.c = na.spline(df$SWC_2_1_1.c,maxgap = gap)
df$SWC_3_1_1.c = na.spline(df$SWC_3_1_1.c,maxgap = gap)

#all soil temps deg C
df$TS_1_1_1.c = na.spline(df$TS_1_1_1.c,maxgap = gap)
df$TS_2_1_1.c = na.spline(df$TS_2_1_1.c,maxgap = gap)
df$TS_3_1_1.c = na.spline(df$TS_3_1_1.c,maxgap = gap)
df$TS_4_1_1.c = na.spline(df$TS_4_1_1.c,maxgap = gap)

#can't gapfill WD as it makes the degrees -50 - 550, and BASE has no u and v components 

```


#Add cardinal directions for wind direction (WD) in BASE df
```{r}
#make sure windDir is numeric 

#(0=north,90=east,180=south,270=west) that the wind is coming from
# northerly wind is 0째, an easterly wind is 90째, a southerly wind is 180째, and a westerly wind is 270째

# Function to assign cardinal directions
df <- df %>%
  mutate(
    Cardinal_Direction = case_when(
      WD >= 337.5 | WD < 22.5  ~ "N",
      WD >= 22.5 & WD < 67.5  ~ "NE",
      WD >= 67.5 & WD < 112.5 ~ "E",
      WD >= 112.5 & WD < 157.5 ~ "SE",
      WD >= 157.5 & WD < 202.5 ~ "S",
      WD >= 202.5 & WD < 247.5 ~ "SW",
      WD >= 247.5 & WD < 292.5 ~ "W",
      WD >= 292.5 & WD < 337.5 ~ "NW",
      TRUE ~ NA_character_ # Handle any missing or invalid values -- not needed in this case 
    )
  )

#checking to see if NA  match up correctly  
sum(is.na(df$WD)) #66561
sum(is.na(df$Cardinal_Direction)) #66561

```


#Save new file of the cleaned met data with minor gapfilling 
```{r}
write.csv(x = df,file = "C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_2016_2023.csv",row.names = F)
```

#Load in and Add ERA5 data
```{r}
#load the ERA5 df from step 0 --> edited to include u and v wind components 

era = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_era5.csv')

df$date = df$ts

summary(era$date) #checking out the timestamps 

#merge cleaned council met df and ERA5 df together
all = merge(era,df,by = 'date',all = T)
all = all[!duplicated(all$date),]
```
#Here we use linear models to make adjustments to the ERA5 data to make it better fit our actual site data and improve later gapfilling


#checking data, and running linear regressions to be able to adjust the ERA5 data to better represent what data from the site is showing*
#the linear reg applies an adjustment to the ERA5 data 
#green data is overwriting the ERA5 pink data -- and we use that linear-reg-adjusted data to help fill in the site data as it's a better representation*

#wind direction in degrees (0-360)
```{r}
#lowercase variables come from ERA; upper-case from our cleaned council met dataset
#KK 10.28.2024: added wind direction 
#from Kyle - these might not agree well as they could come from different height profiles...so while they don't seem to align well at all here, let's not worry about this for now. 

wd.mod = lm(formula = all$WD ~ all$wd)
summary(wd.mod)

all$wd.eramod = all$wd*wd.mod$coefficients[2] + wd.mod$coefficients[1]   

ggplot(data = all)+
  geom_point(aes(date,wd,col='ERA5'))+
  geom_point(aes(date,WD,col='Site'))

ggplot(data = all,aes(wd,WD))+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)
```

#RH - relative humidity 
```{r}
#lowercase variables come from ERA; upper-case from our cleaned council met dataset
#KK 10.28.2024: added the ".c" to use re-cleaned data from Dani 

rh.mod = lm(formula = all$RH.c ~ all$rh)
summary(rh.mod)

all$rh.eramod = all$rh*rh.mod$coefficients[2] + rh.mod$coefficients[1]   

ggplot(data = all)+
  geom_point(aes(date,rh,col='ERA5'))+
  geom_point(aes(date,RH.c,col='Site'))

ggplot(data = all,aes(rh,RH.c))+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)
```
#AirT
```{r}
ggplot(data = all)+
  geom_point(aes(date,airt,col='ERA5'))+
  geom_point(aes(date,all$TA.c,col='Site'))

airt.mod = lm(formula = all$TA.c ~ all$airt)
summary(airt.mod)

all$airt.eramod = all$airt*airt.mod$coefficients[2] + airt.mod$coefficients[1]     

ggplot(data = all,aes(airt,TA.c))+theme_bw()+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)


```



#SoilT
```{r}
ggplot(data = all)+
  geom_point(aes(date,st1,col='ERA5.1'))+
  geom_point(aes(date,st2,col='ERA5.2'))+
  geom_point(aes(date,all$TS_1_1_1.c,col='Site1'))+
  geom_point(aes(date,all$TS_2_1_1.c,col='Site2'))+
  geom_point(aes(date,all$TS_3_1_1.c,col='Site3'))+
   geom_point(aes(date,all$TS_4_1_1.c,col='Site4'))
   
#changed from TS2 to TS3 (tussock) & re-ran code on 6/16/2025
soilt.mod = lm(formula = all$TS_3_1_1.c ~ all$st2)
all$tsoil.eramod = all$st2*soilt.mod$coefficients[2] + soilt.mod$coefficients[1]           

ggplot(data = all)+
  geom_point(aes(date,st2,col='ERA5'))+
  geom_point(aes(date,TS_3_1_1.c,col='Site'))+
  geom_point(aes(date,tsoil.eramod,col='Lin. Reg.'))


# # Fill gaps using the linear model
# df$TS_3_1_1_gapfilled[gaps] <- predict(gapfill_model, 
#                                         newdata = df[gaps, ])
```

#Rg (SW-IN)  - global radiation 
```{r}
# "rad" is from ERA5; SW-IN is from our re-cleaned data, so changed to SW-IN.c to use the re-cleaned data 

ggplot(data = all)+
  geom_point(aes(date,rad,col='ERA5'))+
  geom_point(aes(date,all$SW_IN.c,col='Site'))

ggplot(data = all)+
  geom_point(aes(date,rad,col='ERA5'))+
  geom_point(aes(date,SW_IN.c,col='Site'))+
  scale_x_datetime(limits = as.POSIXct(c('2019-06-20','2019-07-1')))

rad.mod = lm(formula = all$SW_IN.c ~ all$rad)
summary(rad.mod)

all$rad.eramod = all$rad*rad.mod$coefficients[2] + rad.mod$coefficients[1]   

ggplot(data = all,aes(rad,SW_IN.c))+theme_bw()+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)
```


#ws - wind speed 
```{r}
#WS not re-cleaned, so keeping the original WS variable  
ggplot(data = all)+
  geom_point(aes(date,ws,col='ERA5'))+
  geom_point(aes(date,all$WS,col='Site'))

ws.mod = lm(formula = all$WS ~ all$ws)
summary(ws.mod)

all$ws.eramod = all$ws*ws.mod$coefficients[2] + ws.mod$coefficients[1]    

ggplot(data = all,aes(ws,WS))+theme_bw()+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)

```


#Le - latent heat flux 
```{r}
#changed to LE.c to use re-cleaned data 
ggplot(data = all)+
  geom_point(aes(date,le,col='ERA5'))+
  geom_point(aes(date,LE.c,col='Site'))

le.mod = lm(formula = all$LE.c ~ all$le)
summary(le.mod)

all$le.eramod = all$le*le.mod$coefficients[2] + le.mod$coefficients[1]    

ggplot(data = all,aes(le,LE.c))+theme_bw()+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)

```
#H - sensible heat flux 
```{r}
#changed to H.c to use re-cleaned data 
ggplot(data = all)+
  geom_point(aes(date,h,col='ERA5'))+
  geom_point(aes(date,H.c,col='Site'))

h.mod = lm(formula = all$H.c ~ all$h)
summary(h.mod)

all$h.eramod = all$h*h.mod$coefficients[2] + h.mod$coefficients[1]    

ggplot(data = all,aes(h,H.c))+theme_bw()+
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)

```


```{r}
#remove duplicated timestamp rows
df = df[!duplicated(df$ts),]
```


```{r}
#Added ".c" to cleaned data 

#calculate VPD (vapor pressure deficit) from air t and RH in ERA5 df
svp = 610.7*10^((7.5*all$airt)/(237.3+all$airt))
all$vpd.era = ((100 - all$rh)/100)*svp  

#calculate VPD from air t and RH from cleaned council met df (with ".c" to use re-cleaned data)
svp = 610.7*10^((7.5*all$TA.c)/(237.3+all$TA.c))
all$vpd.site = ((100 - all$RH.c)/100)*svp 


ta = ggplot(data = all,aes(airt,TA.c))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 Air T. ("*degree*"C)"))+
  scale_y_continuous(expression("Site Air T. ("*degree*"C)"))

rh = ggplot(data = all,aes(rh,RH.c))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 RH (%)"))+
  scale_y_continuous(expression("Site RH (%)"))

vpd = ggplot(data = all,aes(vpd.era,vpd.site))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 VPD ("*Wm^-2*")"))+
  scale_y_continuous(expression("Site VPD ("*Wm^-2*")"))

#changed from TS2 to TS3 6/16/25 & re-ran code 
st = ggplot(data = all,aes(st2,TS_3_1_1.c))+theme_bw()+
    geom_point(alpha = 0.25)+
    geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 Soil T. ("*degree*"C)"))+
  scale_y_continuous(expression("Site Soil T. ("*degree*"C)"))

sw = ggplot(data = all,aes(rad,SW_IN.c))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 SW ("*Wm^-2*")"))+
  scale_y_continuous(expression("Site SW ("*Wm^-2*")"))

ws = ggplot(data = all,aes(ws,WS))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 WS ("*ms^-1*")"))+
  scale_y_continuous(expression("Site WS ("*ms^-1*")"))

#added wind direction
wd = ggplot(data = all,aes(wd,WD))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 WD ("*ms^-1*")"))+
  scale_y_continuous(expression("Site WD ("*ms^-1*")"))

le = ggplot(data = all,aes(le,LE.c))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 LE ("*Wm^-2*")"))+
  scale_y_continuous(expression("Site LE ("*Wm^-2*")"))

h = ggplot(data = all,aes(h,H.c))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 H ("*Wm^-2*")"))+
  scale_y_continuous(expression("Site H ("*Wm^-2*")"))

#changed from SWC1 to SWC3 6/16/2025
swc = ggplot(data = all,aes(vwc1*100,SWC_3_1_1.c))+theme_bw()+
  geom_point(alpha = 0.25)+
  geom_smooth(method = 'lm')+
  geom_abline(slope = 1,intercept = 0,col='red',lty=2)+
  scale_x_continuous(expression("ERA5 VWC (%)"))+
  scale_y_continuous(expression("Site VWC (%)"))

#Plot agreement between ERA5 and site data 
#png(filename = 'C:/Users/karndt.WHRC/Desktop/sites/YKD/plots/unburnedera5.png',width = 8,height = 8,units = 'in',res = 1500)
plot_grid(ta,rh,vpd,st,sw,ws,wd,le,h,swc)
#dev.off()

#All look decent except for SWC and WD which clearly are having a rough time

```
#Select variables of interest from ERA5 / Reduce down to Council size; add adjusted ERA5 model to final data set
```{r}
#select the data / variables you want from the ERA5 instead of saving all
eras = all[,c('date','rh.eramod','airt.eramod','ws.eramod','wd', 'cardinal_direction','u', 'v','rad.eramod','tsoil.eramod','le.eramod','h.eramod')] #adding the u, v, wd (not wd.eramod since it wasn't aligning well at all), and cardinal_direction 

df = df[!duplicated(df$date),] #delete duplicated dates
df = merge(df,eras,by = 'date',all.x = T) #merge all of those ERA5 variables into our Council dataset 

write.csv(x = df,file = 'C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_2016_2023_era.csv',row.names = F,quote = F)
```




