---
title: "ERA5 WindDir" #Using v and u component from ERA5 (m/s) to translate to 0-360 degrees, and assigning 0-360 degress in both df to a N, E, S, W cardinal coordinates --> testing here, then will add to processing steps 0 and 1 to incorporate the wd in degrees (BASE data has wind direction in 0-360, so this seems the best approach to merge with BASE data)


#( u ) Component: This is the east-west component of the wind. Positive ( u ) values indicate wind blowing from the west towards the east, while negative ( u ) values indicate wind blowing from the east towards the west.
#( v ) Component: This is the north-south component of the wind. Positive ( v ) values indicate wind blowing from the south towards the north, while negative ( v ) values indicate wind blowing from the north towards the south.

output: html_document
date: "2024-11-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Libraries
```{r}
rm(list = ls())

library(data.table)
library(ggplot2)
library(zoo)
library(cowplot)
library(openair)
Sys.setenv(TZ = "UTC")

#set working directory
#setwd("C:/Users/kkent/Documents/Council Data/Council BASE gapfilling")
 
#double check working directory 
getwd()
```

#Load BASE data df from step 1
```{r}
df = fread(input = "C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_2016_2023.csv",na.strings = c('-9999'))#redoing with re-cleaned data 

```


ERA5 data
```{r}
#setwd('C:/Users/karndt.WHRC/Desktop/sites/council/data/')
era = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_era5.csv')

df$date = df$ts

summary(era$date)
summary(df$TIMESTAMP)


#merge together
# all = merge(era,df,by = 'date',all = T)
# all = all[!duplicated(all$date),]
```



#Wind direction from ERA5 from u and v component to 0-360
```{r}
#transform u and v component to 0-360 degrees in ERA5 data 

#make sure u and v are numeric within the ERA dataset 

#- this function converts the u and v components to 0-360 degrees 

#WDIR= (270-atan2(V,U)*180/pi)%360  

windDir <-function(u,v){
  (270-atan2(v,u)*180/pi)%%360 
}


#from NSF you can use either of these two equations, where DperR = (180/pi):
          #Dirgeo = atan2(-U,-V) * DperR <note, -u,-v
                # = 270 - (atan2(V,U) * DperR)


#ERA already has windspeed so don't need to calculate, but here is an equation or see Kyle's code:
#Spd is the horizontal wind speed:  Spd = sqrt(U^2 + V^2)


#wind dir and speed need to be saved as ws and wd for openair to work 

era$wd <-windDir(era$u,era$v)


#in BASE data, it's WS and WD, need to make ws and wd, and make numeric 
df[, 14:15] <- lapply(df[, 14:15], as.numeric)
library(dplyr)
# Rename specific columns
df <- df %>%
  rename(wd = WD, ws = WS)



library(openair)
# type = "month" shows windrose by month 

windRose(era, type = "year")  
windRose(df, type = "year")

#"calm" means a period of time where there is no wind detected / there is still air 

#these two show opposing windRose features but the site data is missing a lot of data 
sum(is.na(df$wd)) #66561 NAs in site data
```

#checking calcs using slightly diff code specific to era df  
```{r}
# Calculate wind direction in degrees

era2 = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling//council_era5.csv')


era2$wd <- (270-atan2(era2$v, era2$u) * 180 / pi + 360) %% 360


era2_filtered <- era2 %>%
  filter(date >= as.Date("2019-06-01") & date <= as.Date("2019-07-01"))



era_filtered <- era %>%
  filter(date >= as.Date("2019-06-01") & date <= as.Date("2019-07-01"))


df_filtered <- df %>%
  filter(date >= as.Date("2019-06-01") & date <= as.Date("2019-07-01"))

# Create the windrose for the filtered data - checking against the other way to calc 
windRose(df_filtered)
windRose(era_filtered)
windRose(era2_filtered)
#era1 and era2 match, feel pretty confident the conversion is correct. But, it does deviate quite a bit from the actual data from the BASE df...*Ask Kyle about thoughts on this, whether to proceed to the step 1 gapfilling etc to prep it to being incorporated into RF 

sum(is.na(df_filtered$wd)) #12 NA's in the data filtered by date....site data and ERA5 don't match up very well 
```
#Checking ERA5 vs BASE wind direction data 
```{r}

#df$date = df$ts

#merge together
all = merge(era,df,by = 'date',all = T)
all = all[!duplicated(all$date),]


# 
# all$date = as.character(all$date)
# 
# all$date= as.POSIXct(all$date, tz="UTC", format = "%Y%m%d%H%M")



library(ggplot2)

# Plotting the data
ggplot() +
  # First dataset (df_filtered)
  geom_point(data = all, aes(x = date, y = wd.x), color = "blue", size = 1) + #ERA data, hourly
  # Second dataset (era_filtered)
  geom_point(data = all, aes(x = date, y = wd.y), color = "red", size = 1) + #BASE data, not gapfilled
  scale_x_datetime(limits = as.POSIXct(c('2019-06-01','2019-07-01')))+
  # Adding labels and theme
  labs(title = "Comparison of Wind Direction - blue = ERA5, red = BASE",
       x = "Date",
       y = "Wind Direction (0-360 degrees)",
       color = "Dataset") +
  theme_minimal()

```

#Assigning cardinal directions to 0-360, 2 ways 
```{r}
#(0=north,90=east,180=south,270=west) that the wind is coming from
# northerly wind is 0°, an easterly wind is 90°, a southerly wind is 180°, and a westerly wind is 270°

# Function to assign cardinal directions
get_cardinal_direction <- function(degree) {
  if (degree >= 337.5 | degree < 22.5) {
    return("N")
  } else if (degree >= 22.5 & degree < 67.5) {
    return("NE")
  } else if (degree >= 67.5 & degree < 112.5) {
    return("E")
  } else if (degree >= 112.5 & degree < 157.5) {
    return("SE")
  } else if (degree >= 157.5 & degree < 202.5) {
    return("S")
  } else if (degree >= 202.5 & degree < 247.5) {
    return("SW")
  } else if (degree >= 247.5 & degree < 292.5) {
    return("W")
  } else if (degree >= 292.5 & degree < 337.5) {
    return("NW")
  }
}

# assign cardinal directions
era$cardinal_direction <- sapply(era$wd, get_cardinal_direction)


#another way: assigning cardinal directions to wind direction in degrees using conditional logic to map ranges of degrees (e.g., 0–360) to specific directions (N, NE, E, etc.). 


library(dplyr)

# Assign cardinal directions based on wind direction in degrees
era <- era %>%
  mutate(
    wind_direction_cardinal = case_when(
      wd >= 337.5 | wd < 22.5  ~ "North",
      wd >= 22.5 & wd < 67.5  ~ "North-East",
      wd >= 67.5 & wd < 112.5 ~ "East",
      wd >= 112.5 & wd < 157.5 ~ "South-East",
      wd >= 157.5 & wd < 202.5 ~ "South",
      wd >= 202.5 & wd < 247.5 ~ "South-West",
      wd >= 247.5 & wd < 292.5 ~ "West",
      wd >= 292.5 & wd < 337.5 ~ "North-West",
      #TRUE ~ NA_character_ # Handle any missing or invalid values -- not needed in this case 
    )
  )
#checking to see if NA values 
sum(is.na(era$wd)) #no NA

```

#assigning cardinal directions based only on u and v to check cardinal directions 

```{r}
#(0=north,90=east,180=south,270=west) that the wind is coming from
#A positive U wind comes from the west, and a negative U wind comes from the east. The V wind component is parallel to the y- axis (i.e. latitude). A positive V wind comes from the south, and a negative V wind comes from the north.

#Note, in this way, true N, E, W, S etc only occur when u or v = 0, which they don't in this dataset, so these will always be NW, NE, SW, SE -> but can help you check out that they relatively match the cardinal assignments from code chunks above  

library(dplyr)

# Assign cardinal directions based on u and v components alone 
era <- era %>%
  mutate(
    wind_direction = case_when(
      u > 0 & v > 0 ~ "SW",  # Positive u and v (wind coming from SW)
      u > 0 & v < 0 ~ "NW",  # Positive u, negative v (wind coming from NW)
      u < 0 & v > 0 ~ "SE",  # Negative u, positive v (wind coming from SE)
      u < 0 & v < 0 ~ "NE",  # Negative u and v (wind coming from NE)
      u == 0 & v > 0 ~ "S",      # u is 0, v positive (from South)
      u == 0 & v < 0 ~ "N",      # u is 0, v negative (from North)
      u > 0 & v == 0 ~ "W",       # u positive, v is 0 (from West)
      u < 0 & v == 0 ~ "E",       # u negative, v is 0 (from East)
      TRUE ~ "Calm"                  # when u = 0 and v = 0
    )
  )

#these give slightly different wind directions than using the degrees ** 


#Testing the calcs - was just double checking various ways to calc cardinal directions 

#example data:
# -2.463795335 u --> neg --> comes from east 
# 5.8896788  v --> pos --> comes from s --> should be a SE wind 

#from NSF you can use either of these two equations where DperR = (180/pi):
          #Dirgeo = atan2(-U,-V) * DperR <note, -u,-v
                # = 270 - (atan2(V,U) * DperR)  <note, this swaps to v,u --> (in the atmo git, this was reversed, which reversed all the caridnal directions / made directions at odds with u,v component directions --> now that order was swapped to V,U, this seems to match the wind dir based on the +/- u and v components*)


wind_direction_test1 <- (atan2(-(-2.463795335), -(5.8896788)) * 180 / pi) %% 360 #where equation has -(u) and -(v), from NSF pg

wind_direction_test2 <- (270-atan2(5.8896788, -2.463795335) * 180 / pi) %% 360 #equation from NSF pg, with 270 - atan... V,U  


print(wind_direction_test1) #157 degrees  --> S/SE --> online wind direction calculator agrees with this 
print(wind_direction_test2) #157 --> now these match, after fixing order of U and V 


# Angle = arctan(V/U) 
#another way--> this is valid as long as U is not equal to zero BUT you also adjust for what quadrant you are in based on whether U and V are positive or negative. The built-in arctan2() function in R will do this for you  
# wind_direction_test3 <- atan(5.8896788/-2.463795335)
# print(wind_direction_test3)   #-1.17, so not right, needs adjustment -- just use arctant2 functions above 







```
#Save new file of ERA5 for next steps --> either do this step within step 0, or after step 1 
#I embedded it into the code in step "0. Council BASE gapfilling with ERA5_github", after making half-hourly & filling in short gaps --> then go on to calc and assign cardinal directions 

```{r}
write.csv(x = eram,file = 'C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_windDir_era5.csv',row.names = F)
```

