---
title: "2.from kyle_Random Forest Council (Neural Network)" 
#Here, we use the Random Forest model to gapfill data based on predictors and real data - it runs simulations and learns patterns, attempting to fill in gaps based on training data and validation data 
#Random Forest is used to do a lot of the gap-filling --we could try to improve it by increasing the trees to allow for more complex relationships and we can clean the data a little more to take out outliers (which Dani did Oct 28, 2024 - we use the re-cleaned version here)

##For Council, NOTE: primary windirection is N-S at this location, S is thermokarst and degradation which can be a big methane source, so wind direction could be important at this site --> NOTE: model is elevating CH4 in the winter, may need to revisit training / other relationships we can use to make sure it's more reasonable (no real winter data available to better train model with)

output: html_document
date: "2024-09-17"
---
#Set working directory
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,dev = 'png')
#knitr::opts_knit$set(root.dir = 'C:/Users/karndt.WHRC/Desktop/sites/council/data') #Edit this to set your working directory
```


#Load in required packages
```{r,error=FALSE,warning=FALSE}
rm(list = ls())
library(data.table)
library(ggplot2)
library(caret)
library(randomForest)
library(Boruta)
#test
Sys.setenv(TZ='UTC')
```

#Load in site data & make training datasets 
```{r,error=FALSE,warning=FALSE}
cf = fread("C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_2016_2023_era.csv") #using the re-cleaned df from Dani 

#subset out by which has complete air t data 
cf = cf[complete.cases(cf$airt.eramod),]


```

#Create subset of data to use for RF training -- use only ERA5 variables for training since RF needs a continuous dataset (no NAs)

```{r}
#create a sub data frame of the variables you want for a neural network (random forest)
#KK Nov 27: added in u, v, wd, and cardinal direction from era df; also changed FC and FCH4 to FC.c, FCH4.c, and SWC_1_1_1.c to use re-cleaned data --> adj to WD from actual data and Cardinal_Direction from actual data as these seem to be era5 data adj by actual data and this uses the actual SWC data from site.
#####


#create subset of the variables wanted for training RF
ncf = data.frame(cf$date,cf$FC.c,cf$FCH4.c,
                 cf$airt.eramod, cf$wd, cf$cardinal_direction, cf$rh.eramod,cf$rad.eramod,
                 cf$ws.eramod,cf$tsoil.eramod,#cf$SWC_1_1_1.c,cf$SWC_2_1_1.c, 
                 cf$SWC_3_1_1.c,
                 cf$h.eramod,cf$le.eramod)
#SWC also broken up by location
#  SWC_1_1_1 % Soil water content (15cm depth) – margin pond
# SWC_2_1_1 % Soil water content (15cm depth) – lichen/berries
# SWC_3_1_1 % Soil water content (15cm depth) - tussock ***** switch to using this one ** (used SWC1 before, by margin pond, switching to SWC3, tussock)

#another way to subset a new df 
# library(dplyr)
# 
# # Select specific variables from 'cf' and store them in 'ncf'
# ncf <- cf %>%
#   select(date, FC.c, FCH4.c, airt.eramod, WD, Cardinal_Direction, rh.eramod, rad.eramod, ws.eramod, tsoil.eramod, SWC_1_1_1.c, h.eramod, le.eramod)  

# View the new dataframe
head(ncf)

#rename for easier names
names(ncf) = c('date','nee',"fch4",'tair', 'wd', 'cardDir','rh','rg','ws','tsoil','swc','h','le')

#make the ERA5 wd numeric 


# #can't have NAs, so need to remove them or use the ERA5 data -- tried this, model still didn't want to run...
# ncf$wd <- na.omit(ncf$wd)
# ncf$cardDir <- na.omit(ncf$cardDir)

#make card dir as factor - 8 levels 
ncf$cardDir <- as.factor(ncf$cardDir)

```


#Calc VPD from air temp and RH
```{r}
#calculate VPD from air t and RH
svp = 610.7*10^((7.5*ncf$tair)/(237.3+ncf$tair))
ncf$vpd = ((100 - ncf$rh)/100)*svp  
```

#create training dataset and testing dataset 
```{r}
#*KK edit 12/27/24 - changing to SWC3 to reflect tussock 

set.seed(123)#sets the start point of models, good for repeatability
cc = ncf[complete.cases(ncf$swc),]#create a gap free data set of the target variable

#use 80% of data set as training set and 20% as test set
sample = sample(c(TRUE, FALSE), nrow(cc), replace=TRUE, prob=c(0.8,0.2))
train  = cc[sample, ]
test   = cc[!sample, ]

#compare to ensure all data sets are representative of total data
hist(cc$swc)
hist(train$swc)
hist(test$swc)

#run random forest to predict missing SWC values - leave out or include wd? included wd here* 

rf.swc = randomForest(formula = swc ~ tair + rh + rg + ws + wd + tsoil + vpd + le + h,data = train,ntree = 150)
#tried various tree numbers, no sig improvement after 300....very minimal diff between 150 and 300, and 150 looks slightly better.

#predict it on the full data set
swc.rf = predict(object = rf.swc,newdata = ncf)
ncf$swc.rf = swc.rf #add the variable to the main data frame


#run random forest to predict missing SWC values - used wd and carDir --> very similar results to when just using wd 
# rf.swc2 = randomForest(formula = swc ~ tair + rh + rg + ws + wd + cardDir + tsoil + vpd + le + h,data = train,ntree = 150)
# #tried various tree numbers, no sig improvement after 300....very minimal diff between 150 and 300, and 150 looks slightly better.
# 
# #predict it on the full data set
# swc.rf2 = predict(object = rf.swc2,newdata = ncf)
# ncf$swc.rf2 = swc.rf2 #add the variable to the main data frame

```


#Check out the SWC gap filling
```{r}
#time series plot of gap filled vs real
ggplot(data = ncf)+
  geom_point(aes(date,swc.rf*100,col='RF'))+
  geom_point(aes(date,swc*100,col='Measured'))+
  scale_y_continuous("SWC (%)")

#validation dataset from only test data
val = merge(test,ncf,by = "date",all.x = T)

ggplot(data = val,aes(swc.rf,swc.x))+theme_bw()+geom_abline(slope = 1,intercept = 0,col='red')+
  geom_point(alpha=0.1)+
  scale_x_continuous(limits = c(0,70),"RF SWC (%)")+
  scale_y_continuous(limits = c(0,70),"Measured SWC (%)")

#summary stats on gap filling
summary(lm(val$swc.x ~ val$swc.rf)) #R2 = 0.97, slope = 1.02

#create final gap free soil moisture
ncf$swc = ifelse(is.na(ncf$swc),ncf$swc.rf,ncf$swc)

#results from using cardinal directions as well as wd in RF --> very similar results, using the version with just wd for simplicity 
#time series plot of gap filled vs real
# ggplot(data = ncf)+
#   geom_point(aes(date,swc.rf2*100,col='RF'))+
#   geom_point(aes(date,swc*100,col='Measured'))+
#   scale_y_continuous("SWC (%)")
# 
# #validation dataset from only test data
# val = merge(test,ncf,by = "date",all.x = T)
# 
# ggplot(data = val,aes(swc.rf2,swc.x))+theme_bw()+geom_abline(slope = 1,intercept = 0,col='red')+
#   geom_point(alpha=0.1)+
#   scale_x_continuous(limits = c(0,70),"RF SWC (%)")+
#   scale_y_continuous(limits = c(0,70),"Measured SWC (%)")
# 
# #summary stats on gap filling
# summary(lm(val$swc.x ~ val$swc.rf2)) #R2 = 0.945

#create final gap free soil moisture
#ncf$swc = ifelse(is.na(ncf$swc),ncf$swc.rf,ncf$swc)
```


#NEE *********************

####Look at range of data 
```{r}
ggplot(data = ncf, aes(x=date, y = nee)) + geom_point()

```


####Create NEE dataset & Examine variable importance
```{r}
set.seed(123)
cc = ncf[complete.cases(ncf$nee),] #complete NEE data only 
cc = subset(cc,cc$nee < 12 & cc$nee > -14) #the range of nee for ncf dataset is -16 - 10 and there are no real major outliers, so this range captures the data well, otherwise use these to set limits on the spread of data to exclude  large outliers and tidy up the model - may need to adjust these limits based on your data*

#this attempts to establish which of the variables have the most important influence on NEE - kyle commented this out, may not be necessary & can take a little while to run but cool to see 

#added in wd (degrees)
boruta = Boruta(nee ~ tair + rh + rg + ws + wd + tsoil + swc + vpd + le + h,data = cc,doTrace = 2,maxRuns = 100)

#plots the variable importance from boruta 
plot(boruta,las = 2)
```

#Let's try to fill NEE with random forest

####Create training and testing dataset 
```{r,error=FALSE,warning=FALSE}
#use 80% of data set as training set and 20% as test set
sample.nee = sample(c(TRUE, FALSE), nrow(cc), replace=TRUE, prob=c(0.8,0.2))
train.nee  = cc[sample.nee, ]
test.nee   = cc[!sample.nee, ]

#Make sure the patterns here look similar so you know the training and testing datasets are representative of the overall dataset 
hist(cc$nee)
hist(train.nee$nee)
hist(test.nee$nee)

#sum(is.na(cc$swc)) #make sure there are no NAs in dataset or RF model won't work 
```

####Determining number of trees for RF 
```{r}
# Train the Random Forest model with a large number of trees
rf_model <- randomForest(nee ~ tair + rh + rg + ws + wd + tsoil + swc + vpd + le + h, data = train.nee, ntree = 1000, importance = TRUE)

# Extract variable importance
importance_values <- importance(rf_model, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf_model, type = 1, main = "Variable Importance (%IncMSE)")

# Evaluate performance with different numbers of trees
error_rates <- rf_model$err.rate[, 1]
plot(error_rates, type = "l", main = "Error Rate vs. Number of Trees", xlab = "Number of Trees", ylab = "Error Rate")
```


```{r}

#original, no wind 
rfnee = randomForest(formula = nee ~ tair + cardDir + rh + rg + ws + tsoil + swc + vpd + le + h,data = train.nee,ntree = 150,importance=T)

pnee = predict(object = rfnee,newdata = ncf)

cf$rfnee = pnee


#using just wd (no cardDir)
rfnee.2 = randomForest(formula = nee ~ tair + cardDir + rh + rg + ws + wd + tsoil + swc + vpd + le + h,data = train.nee,ntree = 150,importance=T)

pnee.2 = predict(object = rfnee.2,newdata = ncf)

cf$rfnee.2 = pnee.2


#trying with just cardinal directions (no wd)
rfnee.3 = randomForest(formula = nee ~ tair + cardDir + rh + rg + ws + cardDir + tsoil + swc + vpd + le + h,data = train.nee,ntree = 150,importance=T)

pnee.3 = predict(object = rfnee.3,newdata = ncf)

cf$rfnee.3 = pnee.3
```


#Checking variable importance using IncMSE
```{r}
#Higher %IncMSE: A higher %IncMSE value means that the variable is more important for the model. Removing this variable would result in a larger increase in the model's prediction error.
#Lower %IncMSE: A lower %IncMSE value means that the variable is less important. Removing this variable would result in a smaller increase in the model's prediction error.


# Train the Random Forest model
rf_model_nee <- randomForest(nee ~ tair + rh + rg + ws + wd + tsoil + swc + vpd + le + h, data = train.nee, importance = TRUE, ntree = 150)

# Extract variable importance
importance_values <- importance(rf_model_nee, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf_model_nee, type = 1, main = "Variable Importance (%IncMSE)")

#Most important variables for NEE are wd (49.81), swc (49.11), ws(41.35)
```

####Time series plots
```{r,error=FALSE,warning=FALSE}
#This overlays the random forest modeled data points with the real data points so you can observe model agreement 

#plot of certain timeframe - no wind 
ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfnee,col='RF'),alpha=0.2)+
  geom_point(aes(date,FC.c,col='Real'),alpha=0.2)+
  scale_x_datetime(limits = as.POSIXct(c("2019-06-15","2019-06-30")))
#plot of all years 
ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfnee,col='RF'),alpha=0.2)+
  geom_point(aes(date,FC.c,col='Real'),alpha=0.2)+
  geom_hline(yintercept = 0)


#plot of certain timeframes using wd (degrees) 
ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfnee.2,col='RF'),alpha=0.2)+
  geom_point(aes(date,FC.c,col='Real'),alpha=0.2)+
  labs(title = "with WD (degrees)")+
  scale_x_datetime(limits = as.POSIXct(c("2019-06-15","2019-06-30")))
#plot of all years 
ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfnee.2,col='RF'),alpha=0.2)+
  geom_point(aes(date,FC.c,col='Real'),alpha=0.2)+
  geom_hline(yintercept = 0)+
    labs(title = "with WD (degrees)")


#plot of certain timeframe - cardinal directions, no wd  
ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfnee.3,col='RF'),alpha=0.2)+
  geom_point(aes(date,FC.c,col='Real'),alpha=0.2)+
  labs(title = "Cardinal Direction (no wd)")+
  scale_x_datetime(limits = as.POSIXct(c("2019-06-15","2019-06-30")))
#plot of all years 
ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfnee.3,col='RF'),alpha=0.2)+
  geom_point(aes(date,FC.c,col='Real'),alpha=0.2)+
  geom_hline(yintercept = 0)+
  labs(title = "Cardinal Direction (no wd)")

#Results very similar, also seems to be predicting slightly lower/more conservative nee compared to site data 
```


####Validation - looking at agreement between modeled data and real data 
```{r,error=FALSE,warning=FALSE}
#changed to FC.c to use re-cleaned variable - original model, no wind direction 
test.data.nee = merge(test.nee,cf,by = 'date',all.x = T)
summary(lm(test.data.nee$FC.c ~ test.data.nee$rfnee)) #R2 = 0.64, slope = 1.02

councilnee = ggplot(data = test.data.nee,aes(rfnee,FC.c))+theme_bw()+
  geom_hline(yintercept = 0,lty=2)+
  geom_vline(xintercept = 0,lty=2)+
  geom_point(alpha=0.2)+
  scale_fill_viridis_c()+
  geom_abline(slope = 1,intercept = 0,col='red',lty=1)+
  scale_x_continuous(limits = c(-20,20),expression('Random Forest NEE ('*mu*mol~CO[2]~m^-2~s^-1*')'))+
  scale_y_continuous(limits = c(-20,20),expression('Eddy Covariance NEE ('*mu*mol~CO[2]~m^-2~s^-1*')'))+
  annotate(geom = 'text',x = 6, y = -8,label=expression(R^2~"= 0.64"),size = 3)+
  annotate(geom = 'text',x = 6,y = -10,label=expression(Slope~"= 1.02"),size = 3)+
  theme(text = element_text(size = 8)) +
  labs(title = "Orig - no wind")

councilnee 

#Added wd (degrees)
test.data.nee = merge(test.nee,cf,by = 'date',all.x = T)
summary(lm(test.data.nee$FC.c ~ test.data.nee$rfnee.2)) #R2 = 0.65, slope = 1.03

councilnee2 = ggplot(data = test.data.nee,aes(rfnee.2,FC.c))+theme_bw()+
  geom_hline(yintercept = 0,lty=2)+
  geom_vline(xintercept = 0,lty=2)+
  geom_point(alpha=0.2)+
  scale_fill_viridis_c()+
  geom_abline(slope = 1,intercept = 0,col='red',lty=1)+
  scale_x_continuous(limits = c(-20,20),expression('Random Forest NEE ('*mu*mol~CO[2]~m^-2~s^-1*')'))+
  scale_y_continuous(limits = c(-20,20),expression('Eddy Covariance NEE ('*mu*mol~CO[2]~m^-2~s^-1*')'))+
  annotate(geom = 'text',x = 6, y = -8,label=expression(R^2~"= 0.65"),size = 3)+
  annotate(geom = 'text',x = 6,y = -10,label=expression(Slope~"= 1.03"),size = 3)+
  theme(text = element_text(size = 8)) +
  labs(title = "with WD (degrees)")

councilnee2 

#used cardinal direction only 
test.data.nee = merge(test.nee,cf,by = 'date',all.x = T)
summary(lm(test.data.nee$FC.c ~ test.data.nee$rfnee.3)) #R2 = 0.65, slope = 1.03

councilnee3 = ggplot(data = test.data.nee,aes(rfnee.3,FC.c))+theme_bw()+
  geom_hline(yintercept = 0,lty=2)+
  geom_vline(xintercept = 0,lty=2)+
  geom_point(alpha=0.2)+
  scale_fill_viridis_c()+
  geom_abline(slope = 1,intercept = 0,col='red',lty=1)+
  scale_x_continuous(limits = c(-20,20),expression('Random Forest NEE ('*mu*mol~CO[2]~m^-2~s^-1*')'))+
  scale_y_continuous(limits = c(-20,20),expression('Eddy Covariance NEE ('*mu*mol~CO[2]~m^-2~s^-1*')'))+
  annotate(geom = 'text',x = 6, y = -8,label=expression(R^2~"= 0.65"),size = 3)+
  annotate(geom = 'text',x = 6,y = -10,label=expression(Slope~"= 1.03"),size = 3)+
  theme(text = element_text(size = 8))+
  labs(title = "with wind cardinal direction")

councilnee3 
```

#Methane 

####Examine variable importance for FCH4
```{r}
set.seed(123)
cc = ncf[complete.cases(ncf$fch4),]

#added wd (degrees)
boruta = Boruta(fch4 ~ tair + rh + rg + ws + wd + tsoil + swc + vpd + le + h,data = cc,doTrace = 2,maxRuns = 100)
boruta$finalDecision #indicates all above are confirmed important variables via a list 
plot(boruta,las = 2) #plots the variables; green = important, yellow = tentative, red = not important 

 

#added wd - cardinal direction
boruta2 = Boruta(fch4 ~ tair + rh + rg + ws + cardDir + tsoil + swc + vpd + le + h,data = cc,doTrace = 2,maxRuns = 100)
plot(boruta2,las = 2)
```


#try random forest for CH4 
```{r,error=FALSE,warning=FALSE}
#KK edit 11.12.2024 - adding trees to see if the model can constrain winter CH4 better

#use 80% of data set as training set and 20% as test set, expanded to try to represent the larger points better

#cc = subset(cc,cc$fch4 < 0.2) #this was kyle trying to get rid of bursts he saw in other datasets that were artificially creating outliers

#essentially we re-run these codes with points of interest, we don't change the predictor variables below (vpd, swc, etc)
cc = subset(cc,cc$fch4 < 50)
cc = subset(cc,cc$fch4 > -25)
sample.ch4 = sample(c(TRUE, FALSE), nrow(cc), replace=TRUE, prob=c(0.8,0.2))
train.ch4  = cc[sample.ch4, ]
test.ch4   = cc[!sample.ch4, ]

hist(cc$fch4)
hist(train.ch4$fch4)
hist(test.ch4$fch4)

summary(cc$fch4)
summary(train.ch4$fch4)
summary(test.ch4$fch4)

#no wind direction 
rfch4 = randomForest(formula = fch4 ~ tair + rh + rg + ws + tsoil + vpd + swc + h + le,data = train.ch4,ntree = 150)

pch4 = predict(object = rfch4,newdata = ncf)

cf$rfch4 = pch4

#added wd (degrees)
rfch4.2 = randomForest(formula = fch4 ~ tair + rh + rg + ws + wd + tsoil + vpd + swc + h + le,data = train.ch4,ntree = 150)

pch4.2 = predict(object = rfch4.2,newdata = ncf)

cf$rfch4.2 = pch4.2

#added cardinal direction (no wd) 
rfch4.3 = randomForest(formula = fch4 ~ tair + rh + rg + ws + cardDir  + tsoil + vpd + swc + h + le,data = train.ch4,ntree = 150)


pch4.3 = predict(object = rfch4.3,newdata = ncf)

cf$rfch4.3 = pch4.3

#added cardinal direction and wd 
rfch4.4 = randomForest(formula = fch4 ~ tair + rh + rg + ws + wd + cardDir + tsoil + vpd + swc + h + le,data = train.ch4,ntree = 150)


pch4.4 = predict(object = rfch4.4,newdata = ncf)

cf$rfch4.4 = pch4.4
```
#IncMSE for CH4 variable importance 
```{r}
#Higher %IncMSE: A higher %IncMSE value means that the variable is more important for the model. Removing this variable would result in a larger increase in the model's prediction error.
#Lower %IncMSE: A lower %IncMSE value means that the variable is less important. Removing this variable would result in a smaller increase in the model's prediction error.


# Train the Random Forest model - same as rfch4.2 model above 
rf_model_CH4 = randomForest(formula = fch4 ~ tair + rh + rg + ws + wd + tsoil + vpd + swc + h + le, data = train.ch4,importance = TRUE, ntree = 150)
# Extract variable importance
importance_values <- importance(rf_model_CH4, type = 1)  # type = 1 for %IncMSE

# Print the importance values
print(importance_values)

# Plot the importance values
varImpPlot(rf_model_CH4, type = 1, main = "Variable Importance (%IncMSE)")

#Most important variables for NEE are swc (81.04), wd (69.06), ws (63.37)
```

#Validation - looking at agreement between modeled data and real data 
## Methane - no wind direction, RF tree = 150 
```{r,error=FALSE,warning=FALSE}
#changed to FCH4.c to use the re-cleaned data - orig model, no wind direction 

test.data.ch4 = merge(test.ch4,cf,by = 'date',all.x = T)

councilch4 = ggplot(data = test.data.ch4,aes(rfch4,FCH4.c))+theme_bw()+
  geom_hline(yintercept = 0,lty=2)+
  geom_vline(xintercept = 0,lty=2)+
  geom_point(alpha=0.2)+
  scale_fill_viridis_c()+
  geom_abline(slope = 1,intercept = 0,col='red',lty=1)+
  scale_x_continuous(limits = c(-50,150),expression('Random Forest '*CH[4]~flux~" ("*mu*mol~CH[4]~m^-2~s^-1*')'))+
  scale_y_continuous(limits = c(-50,150),expression('Eddy Covariance '*CH[4]~flux~" ("*mu*mol~CH[4]~m^-2~s^-1*')'))+
  annotate(geom = 'text',x = 100,y = 20,label = expression(R^2~"= 0.44"),size = 3)+
  annotate(geom= 'text',x = 100,y = 10,label = expression(Slope~"= 1.02"),size = 3)+
  theme(text = element_text(size = 8))

councilch4

summary(lm(test.data.ch4$FCH4.c ~ test.data.ch4$rfch4)) #R2 = 0.44, slope = 1.02 --> R2 before changing tree # was R2=0.52
```
##Methane - adding wind direction 

```{r,error=FALSE,warning=FALSE}

#added wd (degrees)
test.data.ch4.2 = merge(test.ch4,cf,by = 'date',all.x = T)
summary(lm(test.data.ch4.2$FCH4.c ~ test.data.ch4.2$rfch4.2)) #R2 = 0.47, slope = 1.01 

councilch4.2= ggplot(data = test.data.ch4,aes(rfch4.2,FCH4.c))+theme_bw()+
  geom_hline(yintercept = 0,lty=2)+
  geom_vline(xintercept = 0,lty=2)+
  geom_point(alpha=0.2)+
  scale_fill_viridis_c()+
  geom_abline(slope = 1,intercept = 0,col='red',lty=1)+
  scale_x_continuous(limits = c(-50,150),expression('Random Forest FCH4 - with WD ('*mu*mol~CO[2]~m^-2~s^-1*')'))+
  scale_y_continuous(limits = c(-50,150),expression('Eddy Covariance FCH4 ('*mu*mol~CO[2]~m^-2~s^-1*')'))+
  annotate(geom = 'text',x = 50, y = -8,label=expression(R^2~"= 0.47"),size = 3)+
  annotate(geom = 'text',x = 50,y = -15,label=expression(Slope~"= 1.01"),size = 3)+
  theme(text = element_text(size = 8))

councilch4.2 

#used cardinal direction 
test.data.ch4.3 = merge(test.ch4,cf,by = 'date',all.x = T)
summary(lm(test.data.ch4.3$FCH4.c ~ test.data.ch4$rfch4.3)) #R2 =0.46, slope = 0.99

councilch4= ggplot(data = test.data.ch4,aes(rfch42,FCH4.c))+theme_bw()+
  geom_hline(yintercept = 0,lty=2)+
  geom_vline(xintercept = 0,lty=2)+
  geom_point(alpha=0.2)+
  scale_fill_viridis_c()+
  geom_abline(slope = 1,intercept = 0,col='red',lty=1)+
  scale_x_continuous(limits = c(-50,150),expression('Random Forest FCH4 with card Dir ('*mu*mol~CO[2]~m^-2~s^-1*')'))+
  scale_y_continuous(limits = c(-50,150),expression('Eddy Covariance FCH4 ('*mu*mol~CO[2]~m^-2~s^-1*')'))+
  annotate(geom = 'text',x = 6, y = -8,label=expression(R^2~"= 0.46"),size = 3)+
  annotate(geom = 'text',x = 6,y = -10,label=expression(Slope~"= 0.99"),size = 3)+
  theme(text = element_text(size = 8))

councilch4.3 

#used cardinal direction and wd 
test.data.ch4.4 = merge(test.ch4,cf,by = 'date',all.x = T)
summary(lm(test.data.ch4.4$FCH4.c ~ test.data.ch4.4$rfch4.4)) #R2 =0.46, slope = 1.00

councilch4= ggplot(data = test.data.ch4,aes(rfch42,FCH4.c))+theme_bw()+
  geom_hline(yintercept = 0,lty=2)+
  geom_vline(xintercept = 0,lty=2)+
  geom_point(alpha=0.2)+
  scale_fill_viridis_c()+
  geom_abline(slope = 1,intercept = 0,col='red',lty=1)+
  scale_x_continuous(limits = c(-50,150),expression('Random Forest FCH4 with cardDir and WD ('*mu*mol~CO[2]~m^-2~s^-1*')'))+
  scale_y_continuous(limits = c(-50,150),expression('Eddy Covariance FCH4 ('*mu*mol~CO[2]~m^-2~s^-1*')'))+
  annotate(geom = 'text',x = 6, y = -8,label=expression(R^2~"= 0.46"),size = 3)+
  annotate(geom = 'text',x = 6,y = -10,label=expression(Slope~"= 1.00"),size = 3)+
  theme(text = element_text(size = 8))

councilch4.4 

```




#Plots and Validation
```{r,error=FALSE,warning=FALSE}
#changed to FCH4.c to use re-cleaned data 
#Plots modeled vs real data among several dates to observe agreement 

ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfch4.2,col='RF'),alpha=0.5)+
  geom_point(aes(date,FCH4.c,col='Real'),alpha=0.5)+
  labs(title = "FCH4 2019 July 1 - July 30")+
  scale_x_datetime(limits = as.POSIXct(c("2019-07-01","2019-07-30")))
 # scale_y_continuous(limits = c(-0.05,0.05))

ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfch4.2,col='RF'),alpha=0.5)+
  geom_point(aes(date,FCH4.c,col='Real'),alpha=0.5)+
   labs(title = "FCH4 2017")+
  scale_x_datetime(limits = as.POSIXct(c("2017-01-01","2017-12-30")))
 # scale_y_continuous(limits = c(-0.05,0.05))

ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfch4.2,col='RF'),alpha=0.5)+
  geom_point(aes(date,FCH4.c,col='Real'),alpha=0.5)+
  labs(title = "FCH4 2018")+
  scale_x_datetime(limits = as.POSIXct(c("2018-01-01","2018-12-30")))
 # scale_y_continuous(limits = c(-0.05,0.05))

ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfch4.2,col='RF'),alpha=0.5)+
  geom_point(aes(date,FCH4.c,col='Real'),alpha=0.5)+
  labs(title = "FCH4 2019")+
  scale_x_datetime(limits = as.POSIXct(c("2019-01-01","2019-12-30")))
 # scale_y_continuous(limits = c(-0.05,0.05))

ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfch4.2,col='RF'),alpha=0.5)+
  geom_point(aes(date,FCH4.c,col='Real'),alpha=0.5)+
  labs(title = "FCH4 2020")+
  scale_x_datetime(limits = as.POSIXct(c("2020-01-01","2020-12-30")))
 # scale_y_continuous(limits = c(-0.05,0.05))

ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfch4.2,col='RF'),alpha=0.5)+
  geom_point(aes(date,FCH4.c,col='Real'),alpha=0.5)+
  labs(title = "FCH4 2021")+
  scale_x_datetime(limits = as.POSIXct(c("2021-01-01","2021-12-30")))
 # scale_y_continuous(limits = c(-0.05,0.05))

ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfch4.2,col='RF'),alpha=0.5)+
  geom_point(aes(date,FCH4.c,col='Real'),alpha=0.5)+
  labs(title = "FCH4 2022")+
  scale_x_datetime(limits = as.POSIXct(c("2022-01-01","2022-12-30")))
 # scale_y_continuous(limits = c(-0.05,0.05))


ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfch4.2,col='RF'),alpha=0.5)+
  geom_point(aes(date,FCH4.c,col='Real'),alpha=0.5)+
  labs(title = "FCH4 2023")+
  scale_x_datetime(limits = as.POSIXct(c("2023-01-01","2023-12-30")))
 # scale_y_continuous(limits = c(-0.05,0.05))


ggplot(data = cf)+theme_bw()+
  geom_point(aes(date,rfch4.2,col='RF'),alpha=0.2)+
  geom_point(aes(date,FCH4.c,col='Real'),alpha=0.2)


```




#Observing agreement between modeled and real NEE and CH4 flux data 
```{r}
library(cowplot)

#png(filename = 'C:/Users/karndt.WHRC/Desktop/sites/YKD/plots/2023 yk2unburned gapfilling validation.png',width = 6,height = 4,units = 'in',res = 1000)
plot_grid(councilnee2,councilch4.2)
#dev.off()
```
#Rename refnee.2 and fch4.2
```{r}
library(dplyr)
cf <- cf %>%
  rename(rfnee = rfnee.2, rfch4 = rfch4.2)
```


#Save the gapfilled data 
```{r,error=FALSE,warning=FALSE}
#write.csv(x = cf,file = "./council_2017_2023_gf.csv",row.names = F) #_gf to indicate it's been gapfilled 
write.csv(x= cf, file = "C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_2017_2023_gf.csv")
```

