---
title: "4. from kyle_gapfilling and partitioning_Council" #Here, we gapfill and partition the fluxes 
output: html_document
date: "2024-09-17"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load Libraries & set working directory -- need ReddyProc
```{r}
rm(list = ls())
library(REddyProc)


setwd() #use this to set your directory
```

#help("REddyProc-package") - load in the Reddy dataset from previous step 
```{r}
  #+++ Load data with one header and one unit row from (tab-delimited) text file
    EddyData.F <- fLoadTXTIntoDataframe('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/reddy_council_2017_2023.txt')
  # note: use \code{fFilterAttr} to subset rows while keeping the units attributes
```

#subset out problematic repeat rows (may or may not be necessary)
```{r}
#    EddyData.F = EddyData.F[-35088,]  --> not necessary for this df, so commented out 
```
    
#If not provided, calculate VPD from Tair and rH
```{r}
EddyData.F <- cbind(EddyData.F,VPD=fCalcVPDfromRHandTair(EddyData.F$RH, EddyData.F$Tair))
```
  
#Add time stamp in POSIX time format
```{r}
EddyDataWithPosix.F <- fConvertTimeToPosix(EddyData.F, 'YDH', Year='Year', Day='DoY', Hour='Hour')
```

#Initialize R5 reference class sEddyProc for processing of eddy data with all variables needed for processing later
```{r}
 EddyProc.C <- sEddyProc$new('Council', EddyDataWithPosix.F, c('NEE','CH4','H','LE','Ustar','Rg','Tair','Tsoil','RH','VPD'))
  EddyProc.C$sSetLocationInfo(LatDeg =  64.861774, LongDeg = -163.700214, TimeZoneHour = -9) #coordinates for council, with hour difference from UTC (-9)
```

 
#Generate plots of all data in directory \plots (of current R working dir) (Optional)
```{r}
#  EddyProc.C$sPlotDiurnalCycle('CH4')
#  EddyProc.C$sPlotDiurnalCycle('NEE')
#  EddyProc.C$sPlotDiurnalCycle('Rg')
#  EddyProc.C$sPlotDiurnalCycle('Tair')
```

#Plot individual months/years to screen (of current R graphics device)
```{r}
  EddyProc.C$sPlotHHFluxesY('NEE',Year = 2019)
  EddyProc.C$sPlotFingerprintY('NEE', Year=2019)
```

#Fill gaps in variables with MDS gap filling algorithm (without prior ustar filtering) - this ensures correct formatting
```{r}
#we are still using our Random Forest-gap-filled dataset, just run the MDS as well to double check 
#"_f" is created by the MDS gapfilling 
#we gave it a gap-free dataset with random forest, and so the MDS "_f" product should match our gap-filled random forest data
#we do this bc this code looks for very particular variables that MDS provides -- so it's redundant but is what's needed to make sure the data is in the correct format
EddyProc.C$sMDSGapFill('Rg',    FillAll=F) #Fill only the gaps 'Rg'
  EddyProc.C$sMDSGapFill('Tair',  FillAll=F) #Fill only the gaps 'Tair'
  EddyProc.C$sMDSGapFill('VPD',   FillAll=F) #Fill only the gaps 'VPD'
  EddyProc.C$sMDSGapFill('Tsoil', FillAll=F) #Fill only the gaps Tsoil'
  EddyProc.C$sMDSGapFill('RH',    FillAll=F) #Fill only the gaps for RH'
  EddyProc.C$sMDSGapFill('Ustar', FillAll=F) #Fill only the gaps Ustar'
  EddyProc.C$sMDSGapFill('H',     FillAll=F) #Fill only the gaps for H'
  EddyProc.C$sMDSGapFill('LE',    FillAll=F) #Fill only the gaps for 'LE'
  EddyProc.C$sMDSGapFill('NEE',   FillAll=T) #Fill all values to estimate flux uncertainties #can set to F if you need this to run faster
  EddyProc.C$sMDSGapFill('CH4',   FillAll=T)  #Fill all values to estimate flux uncertainties #can set to F if you need this to run faster
```

#Example plots of filled data to screen or to directory \plots (optional)
```{r}
#  EddyProc.C$sPlotFingerprintY('NEE_f',Year.i = 2015)
 # EddyProc.C$sPlotDailySumsY('NEE_f','NEE_fsd', Year=2013) #Plot of sums with uncertainties
  #EddyProc.C$sPlotDailySums('NEE_f','NEE_fsd')
```

#Partition NEE into GPP and respiration (may or may not work based on constraining NEE in step #3)
```{r}
#This attempts another way to constrain the data / partition, but based on constraints set in step #3, it may or may not work if a temperature sensitivity cannot be identified 
  EddyProc.C$sMRFluxPartition()	# night time partitioning -> Reco, GPP
  EddyProc.C$sGLFluxPartition()	# day time partitioning -> Reco_DT, GPP_DT Lasslop, --> this doesn't work, even with some tweaking NEE bounds, move on for now
  
#EddyProc.C$sGLFluxPartition(controlGLPart.l=partGLControl(isBoundLowerNEEUncertainty=FALSE))	# day time partitioning -> Reco_DT, GPP_DT
  

```

#check out data
```{r}
  # plot(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$Rg_f,col='red')
  # points(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$Rg_orig)
  # 
  # plot(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$Tair_f,col='red')
  # points(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$Tair_orig)
  # 
  # plot(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$VPD_f,col='red')
  # points(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$VPD_orig)
  # 
  # plot(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$Tsoil_f,col='red')
  # points(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$Tsoil_orig)
  # 
  # plot(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$LE_f,col='red')
  # points(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$LE_orig)
  # 
  # plot(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$H_f,col='red')
  # points(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$H_orig)
  
  # plot(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$NEE_f,col='red')
  # points(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$NEE_orig)
  # abline(h = 0,col='green')
  # 
  # plot(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$CH4_f,col='red')
  # points(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$CH4_orig)  
  # abline(h = 0,col='green')
  # 

  plot(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$GPP_f,col='green')
  points(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$Reco)
  
  plot(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$GPP_DT,col='green')
  points(EddyProc.C$sTEMP$sDateTime ,EddyProc.C$sTEMP$Reco_DT)
  

```

#test quality of partitioned data
```{r}
#this will not work if constraining NEE in step 3 did not find a temp sensitivity 

 par(mar=c(5,5,5,5))
  plot( EddyProc.C$sTEMP$GPP_DT ~ EddyProc.C$sTEMP$GPP_f); abline(0,1,col='red')
  plot( EddyProc.C$sTEMP$Reco_DT ~ EddyProc.C$sTEMP$Reco); abline(0,1,col='red')
  plot(-EddyProc.C$sTEMP$GPP_DT + EddyProc.C$sTEMP$Reco_DT ~ EddyProc.C$sTEMP$NEE_f ); abline(0,1,col='red')
```

#test results of gapfilling
```{r}
  #  plot(EddyProc.C$sTEMP$NEE_orig,EddyProc.C$sTEMP$NEE_fall);abline(0,1,col='red')
  #  summary(lm(EddyProc.C$sTEMP$NEE_fall ~ EddyProc.C$sTEMP$NEE_orig))
  # # 
  #  plot(EddyProc.C$sTEMP$CH4_orig,EddyProc.C$sTEMP$CH4_fall);abline(0,1,col='red')
  #  summary(lm(EddyProc.C$sTEMP$CH4_fall ~ EddyProc.C$sTEMP$CH4_orig))
```

```{r}
 #names(EddyProc.C$sTEMP)
  # there are some constraints, that might be too strict for some datasets
  # e.g. in the tropics the required temperature range might be too large.
  # It's possible to change these constraints
  #EddyProc.C$sMRFluxPartition(parsE0Regression=list(TempRange.n=2.0, optimAlgorithm="LM")	)  
```

#Example plots of calculated GPP and respiration 
```{r}
#  EddyProc.C$sPlotFingerprintY('GPP_f', Year=2016)
 # EddyProc.C$sPlotHHFluxes('GPP_f')
  #EddyProc.C$sPlotHHFluxesY('GPP_f', Year=2017)
#  EddyProc.C$sPlotHHFluxes('Reco')
```

#Export gap filled and partitioned data to standard data frame
```{r}
FilledEddyData.F <- EddyProc.C$sExportResults()
```

#Save results into (tab-delimited) text file 
```{r}
  CombinedData.F <- cbind(EddyData.F, FilledEddyData.F)

```

# May rename variables to correspond to Ameriflux 
```{r}
  colnames(CombinedDataAmeriflux.F <- renameVariablesInDataframe(CombinedData.F, getBGC05ToAmerifluxVariableNameMapping() ))
  CombinedDataAmeriflux.F$TIMESTAMP_END <- POSIXctToBerkeleyJulianDate( EddyProc.C$sExportData()[[1]] )
  head(tmp <- BerkeleyJulianDateToPOSIXct( CombinedDataAmeriflux.F$TIMESTAMP_END ))
  
  #colnames(tmp <- renameVariablesInDataframe(CombinedData.F, getAmerifluxToBGC05VariableNameMapping() ))
 
 
```

#save as new file
```{r}
 #save as new file 
#for me, this saves it to Github flux folder, not my council data folder 

fWriteDataframeToFile(CombinedData.F, 'council-Results_data_2017_2023.txt', 'data')

```

  