---
title: "6. Ameriflux processing" #another formatting file to put it into Ameriflux organization. Once this is done, we might have to do another clean but should be fairly ready to work with 
output: html_document
date: "2024-09-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/kkent/Documents/Council Data/Council BASE gapfilling') #edit & use this to set your directory
```

#Libraries
```{r}
rm(list = ls())
library(data.table)
library(ggplot2)
library(zoo)

Sys.setenv(TZ = 'UTC')
```

#Load the cleaned, gapfilled Data from last step
```{r}
#df = fread(input = "./council_gapfilled_clean_2019_2023.csv")

df = fread(input = "C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_gapfilled_clean_2017_2023.2.csv")
```

#Format the time stamps at start and end
```{r}
TIMESTAMP_END = df$ts

yeare = format(TIMESTAMP_END,"%Y")
monthe = format(TIMESTAMP_END,"%m")
daye = format(TIMESTAMP_END,"%d")
houre = format(TIMESTAMP_END,"%H")
mine = format(TIMESTAMP_END,"%M")

TIMESTAMP_END = paste(yeare,monthe,daye,houre,mine,sep = '')

TIMESTAMP_START = df$ts - 30*60

years = format(TIMESTAMP_START,"%Y")
months = format(TIMESTAMP_START,"%m")
days = format(TIMESTAMP_START,"%d")
hours = format(TIMESTAMP_START,"%H")
mins = format(TIMESTAMP_START,"%M")

TIMESTAMP_START = paste(years,months,days,hours,mins,sep = '')
rm(daye,days,mine,mins,houre,hours,monthe,months,yeare,years)
```


#Format and name data (based on variables you chose)
```{r}
#makes the names how ameriflux wants -- in this case many variables are probably the same bc we started with an Ameriflux BASE file -- when we add our own data we will have to be careful to make sure the correct names are loaded in

#gas concentrations and variance ################################
 CO2 = df$CO2
 CH4 = df$CH4
 H2O = df$H2O
# 
# CO2_MIXING_RATIO = df$co2_mixing_ratio
# CH4_MIXING_RATIO = ifelse(df$ch4_mixing_ratio<0,NA,df$ch4_mixing_ratio*1000)
# #H2O mole fraction and H2O mixing ratio are identical
# 
# #standard deviation of co2 wet fraction. (Standard deviation is sqrt of the variance)
# CO2_SIGMA = sqrt(df$co2_var)
# CH4_SIGMA = sqrt(df$ch4_var*1000)
# H2O_SIGMA = sqrt(df$h2o_var)
# 
# #Gas Fluxes ######################
 #CO2 flux is umol m-2 s-1     
# FC   = df$FC
# FC_F = ifelse(is.na(df$FC),df$rfnee,df$FC)
# FCH4 = df$FCH4
# FCH4_F = ifelse(is.na(df$FCH4),df$rfch4,df$FCH4)
#FH2O = df$h2o_flux
#______________________________________________________________________________________________________________________
#Changed to FC.c to use the cleaned data KK - 10.29.24 ***********************
FC   = df$FC.c
FC_F = ifelse(is.na(df$FC.c),df$rfnee,df$FC.c)
FCH4 = df$FCH4.c
FCH4_F = ifelse(is.na(df$FCH4.c),df$rfch4,df$FCH4.c)

#GPP_F   = df$GPP_f --> #removing places where we don't have real C flux data and calling that GPP, and the fully gap-filld is GPP_f
GPP     = ifelse(is.na(df$FC.c),NA,df$GPP_f)
GPP_F   = df$GPP_f
RECO     = ifelse(is.na(df$FC.c),NA,df$Reco)
RECO_F = df$Reco

#energy fluxes ###########################
G_1_1_1 = df$G_1_1_1.c
#G_2_1_1 = df$G_2_1_1
#G_3_1_1 = df$G_3_1_1

H       = df$H.c
LE      = df$LE.c

df$TA.c
#meteorology and weather #################################
PA      = df$PA.c #air pressure in kPA
TA      = df$TA.c   # Air T in Deg C
SVP     = 610.7*10^((7.5*TA)/(237.3+TA)) #NOT INCLUDED IN FINAL
RH      = df$RH.c   #RH in %
T_SONIC = df$T_SONIC #sonic temp in Deg C
VPD     = ((100 - RH)/100)*SVP/100 #VPD in hPA
#P       = df$p*1000 # rain in mm

#Radiation #####################################
ALB = df$ALB_1_1_1.c*100 #albedo in % --> this one is new, need to look up 

LW_IN  = df$LW_IN.c  # Wm-2
LW_OUT = df$LW_OUT.c # Wm-2
SW_IN  = df$SW_IN.c   # Wm-2
SW_OUT = df$SW_OUT.c # Wm-2

NETRAD = df$NETRAD.c     # Wm-2

PPFD_IN  = df$PPFD_IN.c   # par in umol m-2 s-1
PPFD_OUT = df$PPFD_OUT.c   # par in umol m-2 s-1

#Soil factors ###################################
#all swc in %
SWC_1_1_1 = df$SWC_1_1_1.c
SWC_2_1_1 = df$SWC_2_1_1.c
SWC_3_1_1 = df$SWC_3_1_1.c


#all soil temps deg C
TS_1_1_1 = df$TS_1_1_1.c
TS_2_1_1 = df$TS_2_1_1.c
TS_3_1_1 = df$TS_3_1_1.c
TS_4_1_1 = df$TS_4_1_1.c
TS_5_1_1 = df$TS_5_1_1 #this one is new, no cleaned version, need to look up what this is (10.29.24*)

#Adding in ERA5 temps to supplement missing site temps 
TA_ERA5 = df$Tair_f
TS_ERA5 = df$Tsoil_f


#wind and turbulence parameters ##############################
df$Tau.c

MO_LENGTH = df$MO_LENGTH
ZL        = df$ZL
TAU       = df$Tau.c

U_SIGMA = df$U_SIGMA
V_SIGMA = df$V_SIGMA
W_SIGMA = df$W_SIGMA

USTAR = df$USTAR

WD     = df$WD
WS     = df$WS
WS_MAX = df$WS_MAX



#From Kyle's YK code: #wind and turbulence parameters ##############################
#stdev is sqr root of variance which is what they are calculating here --> Kyle commented this out, ask why**
# MO_LENGTH = df$L
# ZL        = df$`(z-d)/L`
# TAU       = df$Tau
# 
# U_SIGMA = sqrt(df$u_var)
# V_SIGMA = sqrt(df$v_var)
# W_SIGMA = sqrt(df$w_var)
# 
# USTAR = df$`u*`
# 
# WD     = df$wind_dir
# WS     = df$wind_speed
# WS_MAX = df$max_wind_speed

#Quality flags from EP (Mauder and Foken) ################
# FC_SSITC_TEST   = df$qc_co2_flux
# FCH4_SSITC_TEST = df$qc_ch4_flux
# H_SSITC_TEST    = df$qc_H
# LE_SSITC_TEST   = df$qc_LE
# TAU_SSITC_TEST  = df$qc_Tau
```


Plots to screen / check if look ok 
```{r}
plot(CO2) #umol
plot(CH4) #nmol
plot(H2O) #mmol
```

```{r}
#plot(CO2_MIXING_RATIO) #umol
#plot(CH4_MIXING_RATIO) #nmol
```

```{r}
# plot(CO2_SIGMA)
# plot(CH4_SIGMA) 
# plot(H2O_SIGMA) 
```

#Plotting gapfilled data to screen / make sure it looks ok 
```{r}
#I made them the .c datat above so FC = FC.c, aka the cleaned FC data 
plot(FC) #actual data
plot(FC_F) #"_F" represents gap-filled data
plot(FCH4)
plot(FCH4_F)
#plot(FH2O)
plot(GPP)
plot(GPP_F)
plot(RECO)
plot(RECO_F)


```

```{r}
plot(G_1_1_1)
#plot(G_2_1_1)
#plot(G_3_1_1)
plot(TA_ERA5)
plot(TS_ERA5)
plot(H)
plot(LE)
```


```{r}
plot(PA)
plot(RH)
plot(T_SONIC)
plot(TA)   
plot(VPD) 
#plot(P)
```

```{r}

plot(ALB)
plot(LW_IN)
plot(LW_OUT)
plot(SW_IN)
plot(SW_OUT)
plot(NETRAD)
plot(PPFD_IN)
plot(PPFD_OUT)
```


```{r}
plot(SWC_1_1_1)
plot(SWC_2_1_1)
plot(SWC_3_1_1)


#edit to reflect all temp variables in dataset 
plot(TS_1_1_1)
plot(TS_2_1_1)
plot(TS_3_1_1)
plot(TS_4_1_1)

```

```{r}
plot(MO_LENGTH)
plot(ZL)
plot(TAU)
plot(U_SIGMA)
plot(V_SIGMA)
plot(W_SIGMA)
plot(USTAR)
plot(WD)
plot(WS)
plot(WS_MAX)



```

```{r}
# plot(FC_SSITC_TEST)
# plot(FCH4_SSITC_TEST)
# plot(H_SSITC_TEST)
# plot(LE_SSITC_TEST)
# plot(TAU_SSITC_TEST)
```

#merge everything into a nice and neat file 
```{r}
#checking variables in df to make sure names are present / match


#Create new df with the variables you want -- Council site = US-NGC in Ameriflux 
US_NGC2.2 = data.frame(TIMESTAMP_START,TIMESTAMP_END, #adding ".2" to indicate updated SWC (tussock) - KK 1/3/2025
  CO2,CH4,H2O
#,CO2_MIXING_RATIO,CH4_MIXING_RATIO
#,CO2_SIGMA,CH4_SIGMA ,H2O_SIGMA 
,FC,FCH4#,#FH2O
,FC_F,FCH4_F,GPP,GPP_F,RECO_F,RECO
,G_1_1_1 #,G_2_1_1,G_3_1_1    --> df doesn't have these addiitonal soil temps
,H
,LE
,PA
,RH
,T_SONIC
,TA   
,VPD 
#,P
,ALB
,LW_IN
,LW_OUT
,SW_IN
,SW_OUT
,NETRAD
,PPFD_IN
,PPFD_OUT
,SWC_1_1_1
,SWC_2_1_1
,SWC_3_1_1
,TS_1_1_1
,TS_2_1_1
,TS_3_1_1
,TS_4_1_1
,TS_5_1_1
,TA_ERA5 #added to supplement missing temp data from site 
,TS_ERA5 #added to supplement missing temp data from site 
,MO_LENGTH,ZL,TAU,U_SIGMA,V_SIGMA,W_SIGMA,USTAR,WD,WS,WS_MAX)
#,FC_SSITC_TEST,FCH4_SSITC_TEST,H_SSITC_TEST,LE_SSITC_TEST,TAU_SSITC_TEST)
```

#merge into the site ID -- this is using the old data but for our new council site, we will have a new site ID assigned 
```{r}
siteID = "US-NGC2.2" #adding the 2 to differentiate from US-NGC, as the site will get a new ID, TBD, and ".2" to represent SWC3-tussock dataset
resolution = 'HH'
tsstart = US_NGC2.2$TIMESTAMP_START[1]
tsend   = US_NGC2.2$TIMESTAMP_END[nrow(US_NGC2.2)]

filename = paste(siteID,resolution,tsstart,tsend,sep = '_') #adds the beginning and end time of df to the file name 
filename = paste("./",filename,".csv",sep = "")

write.csv(x = US_NGC2.2, file = filename,row.names = F,na = "-9999",quote = F)

#final doc name will look like this: US-NGC2_HH_201701010000_202309010000.csv --> saves to my Github council folder 

```
