---
title: "Council daily averages" #Some analyses need continuous data, so we calculate daily averages to better organize our visualization and to analyze 
output: html_document
date: "2024-09-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
library(data.table)
library(ggplot2)
library(cowplot)
library(openair)
library(plotrix)
library(signal)
library(svMisc)
library(zoo)
library(stringr)
library(plyr)
library(viridis)
library(lubridate)
library(tidyverse)
library(gridExtra)
library(plotly)
library(RColorBrewer)
library(openair)
```


#read in cleaned and Ameriflux prepped dataset
```{r}
df = fread(input = "C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/US-NGC2_HH_201701010000_202309010000.csv", na.strings="-9999") 


##na.strings = c('-9999','NA','NaN','NAN','-7999')


```

#Create useable timestamp variable 
```{r}
df$TIMESTAMP_END = as.character(df$TIMESTAMP_END)
df$TIMESTAMP_START = as.character(df$TIMESTAMP_START)

df$TIMESTAMP_END = as.POSIXct(df$TIMESTAMP_END, tz="UTC", format = "%Y%m%d%H%M")
df$TIMESTAMP_START = as.POSIXct(df$TIMESTAMP_START, tz="UTC", format = "%Y%m%d%H%M")


```

#Create a new df for daily averages 
```{r}
#using "date" instead of "day" in order to use the openAir and timeAverage packages/functions
df$date = as.Date(df$TIMESTAMP_END)

date = unique(df$date)
df_avg = as.data.frame(date)
```

#Create FC_night in the half-hourly cleaned Ameriflux-prepped dataset from step 6
```{r}
# Filter CO2 Flux data by incoming shortwave values to get nighttime fluxes
df <- df %>%
  mutate(FC_night = ifelse(SW_IN <= 0, FC, NA)) #variable for not gapfilled

df <- df %>%
  mutate(FC_night_F = ifelse(SW_IN <= 0, FC_F, NA)) #variable for gapfilled 
  
```



#Average by day to create continuous daily averages - NOTE: need to make sure to use "date" for openAir and timeAverage code to work*
```{r}
library(openair)
#average entire dataframe with tighter data availability threshold to avoid using large data gaps in calculating the avg -- having at least 50% of the data 
df_avg <- as.data.frame(date)
df_avg <-timeAverage(df, avg.time = "day", data.thresh = 50)


#average dataframe with lighter threshold to get better nighttime data coverage
df_avg_night <- as.data.frame(date)
df_avg_night <-timeAverage(df, avg.time = "day", data.thresh = 10)


#add avg nighttime flux back into the dataset
df_avg$FC_night = df_avg_night$FC_night
df_avg$FC_night_F = df_avg_night$FC_night_F

#From Dani's old code using "day" and a loop to create daily avgs
## average variables by day to create daily average dataframe
# for (i in 1:ncol(df)) {
#   if (class(df[[i]])[1] == 'numeric'){
# 
#   colname = colnames(df)[i]
# 
#   daily_avg_val <- aggregate(df[[colname]] ~ day, data = df, FUN = mean, na.action=na.omit)
# 
#   colnames(daily_avg_val)[2] <- colname
# 
# df_avg = left_join(df_avg, daily_avg_val, by='day')
# 
#   }
#   else{
#     next
#   }
# }


```


#check out the averages to make sure they look reasonable / screen 

```{r}

plot(df_avg$date, df_avg$FC_F)
plot(df_avg$date, df_avg$GPP_F)
plot(df_avg$date, df_avg$RECO)
plot(df_avg$date, df_avg$FC_night_F)

plot(df_avg$TS_3_1_1, df_avg$FC_night_F)

```

# Save Data

```{r}
#these dataframes now include FC_night and FC_night_F 

#half-hourly df with night fluxes 
write.csv(x = df,file = './council_gapfilled_clean_2017_2023_for analysis.csv',quote = F,row.names = F)
#daily avg df with night fluxes 
write.csv(x = df_avg,file = './council_AVG_gapfilled_clean_2017_2023_for analysis.csv',quote = F,row.names = F)

```