---
title: "9. Optional: Gapfilling TA and TS with ERA5" - adj ERA5 to better match site measurements, then using that to gapfill missing soil temp and air temp. Using this for annual and seasonal averages, popssibly to train RF model for CO2 and CH4 predictions - do not use gapfilled for stats, budgets, etc.* 
output: html_document
date: "2025-07-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())

library(data.table)
library(ggplot2)
library(zoo)
library(cowplot)
library(dplyr)
Sys.setenv(TZ = "UTC")
```

#See step 1. Gapfill_met_Council to see how we use linear models to make adjustments to the ERA5 data to make it better fit our actual site data and improve later gapfilling


#checking data, and running linear regressions to be able to adjust the ERA5 data to better represent what data from the site is showing*
#the linear reg applies an adjustment to the ERA5 data 
#green data is overwriting the ERA5 pink data -- and we use that linear-reg-adjusted data to help fill in the site data as it's a better representation*

#Now, we simply create new columns for ERA5-gapfilled air temp and TS_3 (our chosen soil temp probe location)

#load dataset 
```{r}
#using the ".2" version to reflect the updated SWC used in RF - these should have seasons added already to each DOY

#original half-hourly dataframe 
df = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_gapfilled_clean_2017_2023_for analysis.2.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))

#needs continuous day variable - daily avg, calculated in processing steps
df_avg = fread('C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_AVG_gapfilled_clean_2017_2023_for analysis.2.csv',na.strings = c('-9999','NA','NaN','NAN','-7999'))


```

#For reference, what we did in step 1 

#Check the agreement between measured temps & ERA5 temps, use linear reg model to adjust ERA5 to more closely resemble the site's measurements & use the modified version to gapfill 

#AirT
```{r}
# ggplot(data = all)+
#   geom_point(aes(date,airt,col='ERA5'))+
#   geom_point(aes(date,all$TA.c,col='Site'))
# 
# airt.mod = lm(formula = all$TA.c ~ all$airt)
# summary(airt.mod) #R2 = 0.96
# 
# all$airt.eramod = all$airt*airt.mod$coefficients[2] + airt.mod$coefficients[1]     
# 
# ggplot(data = all,aes(airt,TA.c))+theme_bw()+
#   geom_point()+
#   geom_smooth(method = 'lm')+
#   geom_abline(slope = 1,intercept = 0,col='red',lty=2)
```

#SoilT
```{r}
# ggplot(data = all)+
#   geom_point(aes(date,st1,col='ERA5.1'))+
#   geom_point(aes(date,st2,col='ERA5.2'))+
#   geom_point(aes(date,all$TS_1_1_1.c,col='Site1'))+
#   geom_point(aes(date,all$TS_2_1_1.c,col='Site2'))+
#   geom_point(aes(date,all$TS_3_1_1.c,col='Site3'))+
#    geom_point(aes(date,all$TS_4_1_1.c,col='Site4'))
#    
# #changed from TS2 to TS3 (tussock) & re-ran code on 6/16/2025
# soilt.mod = lm(formula = all$TS_3_1_1.c ~ all$st2)
# summary(soilt.mod) #R2 = 0.87
# all$tsoil.eramod = all$st2*soilt.mod$coefficients[2] + soilt.mod$coefficients[1]           
# 
# ggplot(data = all)+
#   geom_point(aes(date,st2,col='ERA5'))+
#   geom_point(aes(date,TS_3_1_1.c,col='Site'))+
#   geom_point(aes(date,tsoil.eramod,col='Lin. Reg.'))
```


#Create a new column for gapfilled air temp and TS_3 soil temp 

#Soil temp TS_3 - HH
```{r}
# Create a copy of observed soil temp, rename as gapfilled 
df$TS_3_gapfilled <- df$TS_3_1_1

# Identify gaps (missing observed data but available ERA5 data)
gaps <- is.na(df$TS_3_1_1) & !is.na(df$TS_ERA5)

# gapfill with ERA5-modified values, keep orig values, add in era.mod values where there are missing values 
df$TS_3_gapfilled = ifelse(is.na(df$TS_3_1_1), df$TS_ERA5, df$TS_3_1_1)

# Add a flag to track which values were gap-filled
df$TS_3_gapfilled_flag <- ifelse(gaps, "gapfilled", "observed")

summary(df$TS_3_gapfilled)

```


#Soil temp TS_3 - Daily Avg 
```{r}
# Create a copy of observed soil temp, rename as gapfilled 
df_avg$TS_3_gapfilled <- df_avg$TS_3_1_1

# Identify gaps (missing observed data but available ERA5 data)
gaps <- is.na(df_avg$TS_3_1_1) & !is.na(df_avg$TS_ERA5)

# gapfill with ERA5-modified values, keep orig values, add in era.mod values where there are missing values 
df_avg$TS_3_gapfilled = ifelse(is.na(df_avg$TS_3_1_1), df_avg$TS_ERA5, df_avg$TS_3_1_1)

# Add a flag to track which values were gap-filled
df_avg$TS_3_gapfilled_flag <- ifelse(gaps, "gapfilled", "observed")

summary(df_avg$TS_3_gapfilled)

```



#Air temp - HH
```{r}
# Create a copy of observed soil temp, rename as gapfilled 
df$TA_gapfilled <- df$TA

# Identify gaps (missing observed data but available ERA5 data)
gaps <- is.na(df$TA) & !is.na(df$TA_ERA5)

# gapfill with ERA5-modified values, keep orig values, add in era.mod values where there are missing values 
df$TA_gapfilled = ifelse(is.na(df$TA), df$TA_ERA5, df$TA)

# Add a flag to track which values were gap-filled
df$TA_gapfilled_flag <- ifelse(gaps, "gapfilled", "observed")

summary(df$TA_gapfilled)

```


#Air Temp - Daily Avg 
```{r}
# Create a copy of observed soil temp, rename as gapfilled 
df_avg$TA_gapfilled <- df_avg$TA

# Identify gaps (missing observed data but available ERA5 data)
gaps <- is.na(df_avg$TA) & !is.na(df_avg$TA_ERA5)

# gapfill with ERA5-modified values, keep orig values, add in era.mod values where there are missing values 
df_avg$TA_gapfilled = ifelse(is.na(df_avg$TA), df_avg$TA_ERA5, df_avg$TA)

# Add a flag to track which values were gap-filled
df_avg$TA_gapfilled_flag <- ifelse(gaps, "gapfilled", "observed")

summary(df_avg$TA_gapfilled)

```

#Save file with gapfilled columns and gf flags 
```{r}
#half-hourly - ".2" to indicated from updated SWC, ".3" to indicate added temp gf columns 
write.csv(x = df,file = 'C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_gapfilled_clean_2017_2023_for analysis.3.csv',quote = F,row.names = F)

#daily avg df - ".2" to indicated from updated SWC. ".3" to indicate added temp gf columns 
write.csv(x = df_avg,file = 'C:/Users/kkent/Documents/Council Data/Council BASE gapfilling/council_AVG_gapfilled_clean_2017_2023_for analysis.3.csv',quote = F,row.names = F)

```


#Noting % of each year that is gapfilled vs observed - HH 
```{r}
# Load dplyr if not already loaded
library(dplyr)

# Calculate percentages for Air Temperature (TA)
TA_summary_HH <- df %>%
  group_by(year) %>%
  summarise(
    total_records = n(),
    observed_count = sum(TA_gapfilled_flag == "observed", na.rm = TRUE),
    gapfilled_count = sum(TA_gapfilled_flag == "gapfilled", na.rm = TRUE),
    missing_count = sum(is.na(TA_gapfilled)),
    observed_percent = round((observed_count / total_records) * 100, 1),
    gapfilled_percent = round((gapfilled_count / total_records) * 100, 1),
    missing_percent = round((missing_count / total_records) * 100, 1)
  ) %>%
  mutate(variable = "Air_Temperature")

# Calculate percentages for Soil Temperature (TS_3_1_1)

TS_summary_HH <- df %>%
  group_by(year) %>%
  summarise(
    total_records = n(),
    observed_count = sum(TS_3_gapfilled_flag == "observed", na.rm = TRUE),
    gapfilled_count = sum(TS_3_gapfilled_flag == "gapfilled", na.rm = TRUE),
    missing_count = sum(is.na(TS_3_gapfilled)),
    observed_percent = round((observed_count / total_records) * 100, 1),
    gapfilled_percent = round((gapfilled_count / total_records) * 100, 1),
    missing_percent = round((missing_count / total_records) * 100, 1)
  ) %>%
  mutate(variable = "Soil_Temperature")
```

#Noting % of each year that is gapfilled vs observed using df_avg
```{r}
# Load dplyr if not already loaded
library(dplyr)

# Calculate percentages for Air Temperature (TA)
TA_summary <- df_avg %>%
  group_by(year) %>%
  summarise(
    total_records = n(),
    observed_count = sum(TA_gapfilled_flag == "observed", na.rm = TRUE),
    gapfilled_count = sum(TA_gapfilled_flag == "gapfilled", na.rm = TRUE),
    missing_count = sum(is.na(TA_gapfilled)),
    observed_percent = round((observed_count / total_records) * 100, 1),
    gapfilled_percent = round((gapfilled_count / total_records) * 100, 1),
    missing_percent = round((missing_count / total_records) * 100, 1)
  ) %>%
  mutate(variable = "Air_Temperature")

# Calculate percentages for Soil Temperature (TS_3_1_1)

TS_summary <- df_avg %>%
  group_by(year) %>%
  summarise(
    total_records = n(),
    observed_count = sum(TS_3_gapfilled_flag == "observed", na.rm = TRUE),
    gapfilled_count = sum(TS_3_gapfilled_flag == "gapfilled", na.rm = TRUE),
    missing_count = sum(is.na(TS_3_gapfilled)),
    observed_percent = round((observed_count / total_records) * 100, 1),
    gapfilled_percent = round((gapfilled_count / total_records) * 100, 1),
    missing_percent = round((missing_count / total_records) * 100, 1)
  ) %>%
  mutate(variable = "Soil_Temperature")

# Combine both summaries
combined_summary <- rbind(TA_summary, TS_summary)

# Print the results
print("Air Temperature Gap-filling Summary by Year:")
print(TA_summary)

print("Soil Temperature Gap-filling Summary by Year:")
print(TS_summary)

# Optional: Create a nice formatted table
library(knitr)
kable(combined_summary, caption = "Gap-filling Summary by Year and Temp Variable")
```


